---
title: "Function for Indirect N2O"
author: "Dr. Chih-Yu Hung"
date: "2024-05-01"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Indirect N2O emissions

Canada's National Inventory Report estimate indirect N2O emissions from volatilization and leaching/runoff. 

### 1.Volatilization
The volatilization was estimated based on a series of coefficent that influence NH3 volatilization. The coefficients were developed by Bouwman (2002) and then valudated by Sheppard et al.(2010), with Canada's data. The NH3 is converted to N2O by IPCC emission factor. 

### 2. Leaching/runoff
The leaching/runoff may cause NH3 emissions and then being converted to N2O emissions. Canada's methodlogies estimate leaching N proportion with the ratio of precipitation/ potential evapotranspiration. 


```{r volatilization}
N2OEF_volatilization <- function(SiteData) {
  
#Check NA   
  if (any(is.na(SiteData))) {
  print("There are missing values (NA) in the data.")
  stop()  
  }
  if (any(SiteData < 0)) {
  print("There are negative values in the data.")
  stop()
  } 
  
  
#Example for SiteData
#   RegionID, suggest using HUP
#   Year
#   Cropy tpe, Annual or Perennial
#   Fertilizer type, Urea, Urea ammonia, Anhydrous ammonia, and Other
#   Application method, Broadcast and Incorporated
#   Soil chemical, soil pH (numeric or characteristic Low and High), soil CEC (numeric or characteristic Low and High)
#   Climate

  
    
#assign the variables
   Precip <- SiteData[, grepl("preci|^[Pp]$", names(SiteData),ignore.case = TRUE)]
   Evapo  <- SiteData[, grepl("evapotranspiration|potentail eva|PE|PET", names(SiteData),ignore.case = TRUE)]
   Topo   <- SiteData[, grepl("topo", names(SiteData),ignore.case = TRUE)]
   Crop   <- SiteData[, grepl("crop", names(SiteData),ignore.case = TRUE)]
   NSE    <- SiteData[, grepl("non-growing|growing|freeze-thaw|FTC|NSE", names(SiteData),ignore.case = TRUE)]
   NS     <- SiteData[, grepl("NS|source|^[Nn]$", names(SiteData),ignore.case = TRUE)]
   Frac_C <- SiteData[, grepl("Coarse", names(SiteData),ignore.case = TRUE)]
   Frac_M <- SiteData[, grepl("Medium", names(SiteData),ignore.case = TRUE)]
   Frac_F <- SiteData[, grepl("Fine", names(SiteData),ignore.case = TRUE)]
   
#Check the numbers
#Precipitation (mm), in case there is a wrong unit (e.g. cm)   
   if (any(Precip < 80)) {
    print ("There are regions with low growing season precipitation (<80 mm) or unit is not correct, please check")
   }
   if (any(Precip > 1000)) {
    print ("There are regions with high growing season precipitation (> 1000 mm), please check")  
   } 

#Check topography
   if (any(Topo > 1)) {
    stop("Topography are not between 0 and 1")
   }
   
#Check the cropping system   
   if (is.numeric(Crop)){
     if(any(Crop > 4))
     stop("Cropping system codes should be between 1 to 4") #1 annual, 2 perennial, 3 impat, 4 unimpast 
   } 
   
   if (is.character(Crop)){
     allowed_values <- c("annual", "perennial", "impast", "unimpast")
     # Check if all elements in the vector are among the allowed values
    if (!all(tolower(Crop) %in% allowed_values)) {
    stop("Allowed values are 'annual', 'perennial', 'impast', 'unimpast'.")
      }
    }

#Check the Nitrogen source (NS)
   if (is.numeric(NS)){
     if(any(NS > 4))
    stop("Nitrogen source code should be between 1 to 4") #1 inorganic, 2 animal manure, 3 Crop residue, 4 biosolid    
   }
   if (is.character(NS)) {
     NS <- toupper(NS)
     NS <- ifelse(NS %ini% c("FSN","IN","SYNTHETIC Fertilizer", "INORGANIC FERTILZER"), 1,
           ifelse(NS %in% c("AM","ANIMAL MANURE"), 2,
           ifelse(NS %in% c("CROP RESIDUE","CROP","RESIDUE"), 3,
           ifelse(NS %in% c("BIOSOLID"), 4))),
           NS)
   if(any(is.character(NS))){
       stop("Unidentified N source")
     }
   }
   
#Check soil texture faction
  if (!all(sapply(list(Frac_C, Frac_M, Frac_C), is.numeric))) {
  stop("Not all vectors numeric in soil texture fraction")
    }
  # Check if vectors are between 0 and 1
  if (!all(Frac_C >= 0 & Frac_C <= 1) || 
      !all(Frac_M >= 0 & Frac_M <= 1) || 
      !all(Frac_F >= 0 & Frac_F <= 1)) {
    stop("Fraction of each soil texture must be between 0 and 1.")
  } 
    if (all(identical((Frac_C + Frac_M + Frac_F),1))) { 
      stop ("Sum of Frach of soil texture must be 1")
      }

#Calculate the base emission factor applied with texture ratio factor in a region
Result <- data.frame()

Result$Region <-SiteData[,1]
Result$Year <- SiteData[,2]
      
Result$EF_base <- ifelse(Precip>Evapo,exp(0.00558*Precip-7.7),
                           exp(0.00558*Evapo-7.7)*Topo + exp(0.00558*P-7.7)*(1-Topo))
#Calculate the weighted ratio factor for soil texutre
Result$Wtd_RF_TX <- (Frac_C*0.49 + Frac_M*1 + Frac_F*2.55)

#Introduce the Weighted ratio factor to calculate Base emission factor
Result$EF_base <- EF_base * Wtd_RF_TX

#Convert NSE, NS, and cropping system to their ratio factors
NSE <- ifelse(NSE == 1, 1/0.634, 1)
Crop <- ifelse(Crop == 1 | Crop == "annual", 1, 0.19)
NS[Crop == 1 & NS == 1] <- 1
NS <- ifelse(NS == 1, NS, 0.84)
#Calculate the EF 
Result$EF <- EF_base * NSE * Crop * NS
   
}
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
