---
title: "N and Crop data preparation"
author: "Dr. Chih-Yu Hung"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Prepare Real fertilization amount for each crop in Canada

1. `FF2_Fertilizer_N_by_crop` provides the real fertilizer usage after justification (recommended fertilization, manure application, and provicinal sell amount)
2. `INPUT-Crop-TYPEIII` provides the crop ID used by AAFC. 

### Target crop
Barley,
Camelina #not in CropID
Canola
Corn (grain) #keep silage in case
Field peas
Rye (excluding spring rye) #keep spring in case
Sorghum #not in CropID
Soybeans
Triticale
Wheat (durum)
Wheat (spring and winter). 


```{r fertilizer for ecodistrcit, echo=FALSE}
Fert_crop <- read.csv("Input/FF2_Fertilizer_N_by_crop.csv") # kg
CropID <-read.csv("Input/INPUT-Crop-TypeIII.csv") #T

#select the crop I need
CropID <- CropID %>%
  #filter (CropIII_ID %in% c(21,22,18,14,15,11,3,4,19,20,24,27,16))
  filter (Crop_TypeIII_Name %in% c("Barley","Canola","Grain Corn","Silage Corn","Dry Pea","Spring Rye","Fall Rye","Soy bean","Triticale","Durum wheat","Spring wheat","Winter wheat"))

Fert_crop <-Fert_crop %>%
  filter (CropIII_ID %in% CropID$CropIII_ID) %>%
  filter(Inv_Year_ID >= 2016)

#save this table
#save(Fert_crop, file = "Input/Crop type and real Fertilizer_2016-2022.RData")
```


## Input the crop production from Stat Canada
```{r Crop production, ech0=FALSE}
Production <- read.csv("Input/Crop_production_2012-2022.csv") # metric ton for production
#Wrangle the data
Production[is.na(Production)] <- 0
# replace province name 
province_mapping <- data.frame(
  full_name = sort(unique(Production$Province)),
  abbreviation = sort(unique(Fert_crop$Province_ID))
)

#replace province full name with abbreviation
Production$Province <- province_mapping$abbreviation[match(Production$Province, province_mapping$full_name)]

#Cannot find a proper way to replace the Crop, This is a temporary way
#Assign CropID for crops
Production <- Production %>%
  mutate(CropID = case_when(
    Crop == "Canola (rapeseed)" ~ 3,
    Crop == "Soybeans" ~ 4,
    Crop == "Peas, dry" ~ 11,
    Crop == "Wheat, spring" ~ 14,
    Crop == "Wheat, winter remaining"~ 15,
    Crop == "Wheat, durum" ~ 16,
    Crop == "Barley" ~ 18,
    Crop == "Rye, spring" ~ 19,
    Crop == "Rye, fall remaining" ~ 20,
    Crop == "Corn for grain" ~ 21,
    Crop == "Corn for silage" ~ 22,
    Crop == "Triticale" ~ 24,
    TRUE ~ 100)) %>%
    filter (Year >= 2016)  
  
#save crop production
#save(Production, file= "Input/Crop production_prov_2016-2022.RData")
```