---
title: "Volatilization EF"
author: "Dr. Chih-Yu Hung"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Data source

read the file A#_Ammonia_EF_annual_by_Prov
This is table from 
"D:\1_InventoryContributions\1_Agriculture\2024\Data\Processing\2024_DB\5-Ammonia\NH3_NAESI_Synt_Fert.accdb"
I'm using this to have the total NH3 emissions and applied fertilizer and IEF
The IEF can be convert to N2O by applying EF4

```{r Data}
IEF_NH3 <- read.csv("../Input/A3_Ammonia_EF_Annual_by_Prov.csv")
EF4 <- read.csv("../Input/EF4.csv") #Ecod, Ecozone, Prov

#Assign EF4 for ecozones.
low.Ecozone <- c("Taiga Plain","Boreal Plains","Prairie","Montane Cordillera")
high.Ecozone <- c("Boreal Shield","Atlantic Maritime","MixedWood Plain","Pacific Maritime")
EF4 <- EF4 %>%
  mutate(EF4 = case_when(
    ZONE_NAME %in% low.Ecozone ~ 0.005,
    ZONE_NAME %in% high.Ecozone ~ 0.014,
    TRUE ~ 99
  )
  )

#Average EF4 for province
EF4.prov <- as.data.frame(tapply(EF4$EF4, EF4$Province_ID, mean))
colnames(EF4.prov) <-"Avg_EF4"
EF4.prov <- rownames_to_column(EF4.prov, var = "Province_ID")




#Obtain IEF for N2O
IEF_N2O <-IEF_NH3 %>%
  inner_join(EF4.prov,by = c("Province_ID" = "Province_ID")) %>%
  mutate(IEF_N2O = IEF * Avg_EF4)
```


## Calculate the weighted IEF for every province

```{r fertilization allocation}
IEF_NH3[is.na(IEF_NH3)] <- 0
IEF_NH3_wtd <- IEF_NH3 %>%
  group_by(Province_ID, Year) %>%
  summarise(wtd_IEF = ifelse(sum(Fertilizer_Applied) >0, sum(Emissions)/sum(Fertilizer_Applied),0))

#Obtain IEF for N2O
IEF_N2O_wtd <-IEF_NH3_wtd %>%
  inner_join(EF4.prov,by = c("Province_ID" = "Province_ID")) %>%
  mutate(IEF_N2O_wtd = wtd_IEF * Avg_EF4)  
  

load("../Input/Crop type and real Fertilizer.RData") #Fert_crop

Fert_crop_prov <- Fert_crop %>%
  group_by(Province_ID, CropIII_ID) %>%
  summarise(Crop_Real_N = sum(Crop_Real_N))


#inner join the weighted IEF to crop and fertilizer in ecodistrict
Fert_crop_IEF <- IEF_N2O_wtd %>%
  filter(Year == 2022) %>%
  inner_join(Fert_crop_prov, by = c("Province_ID" = "Province_ID"))

#Calculate the total N2O emissions for crops in a province
N2O_vola_crop <- Fert_crop_IEF %>%
  mutate(N2O_vola = Crop_Real_N * IEF_N2O_wtd)

#read production data and calculate the N2O intensity
load("../Input/Crop_production_prov.RData") #Production
N2O_vola <- N2O_vola_crop %>%
  inner_join(Production, by = c("Province_ID" = "Province", 
                                "CropIII_ID" = "CropID")) %>%
  mutate(N2O_intensity = ifelse(Production >0, N2O_vola/(Production*10^3), 0)) %>%
  select(Province_ID,CropIII_ID,Crop,N2O_intensity)

#combine the region
MT <- c("NL","NS","PE","NB")
N2O_vola_region <- N2O_vola %>%
  mutate (Region = case_when(
          Province_ID %in% MT ~ "MT",
          TRUE ~Province_ID)) %>%
  group_by(Region,CropIII_ID,Crop) %>%
  summarise(EF = round(mean(N2O_intensity),5))

# Convert long table to wide table
N2O_vola_result <- pivot_wider(N2O_vola_region,
                              names_from = Region, values_from = EF)
write.csv(N2O_vola_result, "../Output/Indirect N2O_volatilization.csv")

```

```{r Calculation and Preparation for tables, echo=FALSE}
#Take the average
IEF_N2O_prov <- IEF_N2O %>%
  group_by(Year,Province_ID) %>%
  summarise(Total_Fert = sum(Fertilizer_Applied,na.rm = TRUE),
            Total_Emissions = sum(Emissions, na.rm= TRUE),
            Avg_EF4 = mean(Avg_EF4)) %>%
  mutate(IEF_NH3 = Total_Emissions / Total_Fert,
         IEF_N2O = IEF_NH3*Avg_EF4)

#Make a table for region, i.e. put NL, NS, PE, and NB to Maritime
MT <- c("NL","NS","PE","NB")
IEF_N2O_region <- IEF_N2O_prov %>%
  mutate(Region = case_when(
    Province_ID %in% MT ~ "MT",
    TRUE ~Province_ID)) %>%
  group_by(Year,Region) %>%
  summarise(Total_Fert = sum(Total_Fert, na.rm = TRUE),
            Total_Emissions = sum(Total_Emissions, na.rm = TRUE),
            Avg_EF4 = mean(Avg_EF4)) %>%
  mutate(IEF_NH3 = Total_Emissions/Total_Fert,
         IEF_N2O = IEF_NH3*Avg_EF4)

#Calculate the GWP for the IEF, AR5 numbers
IEF_N2O_region_GWP <- IEF_N2O_region %>%
  mutate(CO2e = IEF_N2O *44/28 *265)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
