---
title: "Direct N2O emissions from crop residue"
author: "Dr. Chih-Yu Hung"
date: "2025-06-25"
output: html_document
---

```{r setup, include=FALSE}
source("../Setup.R")
```

## Data descrpition
To calculate direct N2O emissions from crop residue for regionalized crop, we need
1. Precipitation and Potential evaporation (INPUT-EF_Fixed_Para)
2. Topography (INPUT-EF_Fixed_Para)
3. Fraction of soil texture (INPUT-EF_dep_Para_texture)


## Calculate the BaseEF and EF

```{r EF caluclation}
source("Function for N2O EF_Direct.R")

SiteData<-read_rds("../Input/CA_Direct_Crop_2016-2022.rds") #SiteData
#SiteData<-read_rds("../Input/CA_Direct_Crop_2016-2022_WO_Silage.rds") # Without silage

#The last 5 years
SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))

#Calculate the EF
Result <- N2OEF_direct(SiteData) 

#Atlantic provinces
AT <- c("NL","NS","PE","NB")

N2O_long<- Result[[1]] %>%
  mutate (Region = case_when(         #combine the region
          Province %in% AT ~ "AT",
          TRUE ~ Province)) %>%
  group_by(Region,Crop) %>%
  summarise(Tot.Fert = sum(Tot.Fert), Tot.N2O = sum(Tot.N2O)) %>%
  mutate(N2O.IEF =  round(ifelse(Tot.Fert >0, Tot.N2O/Tot.Fert,0),8))

```


## Add Sorghum, camelina and triticale data
In Canada, we have very little amount of grain sorghum. Considering the close relationship between grain Sorghum and corn and their similar N requirement and fertilizer management plan. We are going to use N2O EF of corn for sorghum. 

In Canada, the optimal fertilization plan for camelina was not determined. Recommended application range is 60-100 kg N /ha. However, CFIA (2017) suggested using fertilization plan of canola if further information is not available. 

Other province didn't have fertilizer guide for cameline (e.g. Manitoba Agriculture), AFOLU therefore used EFs of canola.  

Only QC, ON, SK, MB, AB (QC, ON and MB has very small amount <10000 metric ton) have triticale production


Ref:
Jordan B., Fertilizing to help maximize grain sorghum yields. https://www.cropscience.bayer.us/articles/channel/fertilizing-to-help-maximize-grain-sorghum-yields. access 24 May, 2024.
Vanderlip, R., Roozeboom, K., Fjell, D., Shroyer, J., Kok, H., Regehr, D., Whitney, D., Rogers, D.H., Alam, M., Jardine, D. Brooks, H.L., Taylor, R.K., Harner III, J.P., and Langemeier, L.N. 1998. Grain sorghum production handbook. C-687 (Revised). Kansas State University. https://bookstore.ksre.ksu.edu/pubs/c687.pdf.
Haag, L. (In cooperation with: Diaz, D. R., Obour, A., and Schlegel, A.) 2018. Grain sorghum fertility management. Kansas State University. https://www.northwest.k-state.edu/agronomy/documents/presentation_pdfs/2018%20Grain%20Sorghum%20Fertility%20Management%20SorghumU%202spp.pdf.
CFIA (2017) The Biology of Camelina sativa (L.) Crantz (Camelina). https://inspection.canada.ca/en/plant-varieties/plants-novel-traits/applicants/directive-94-08/biology-documents/camelina-sativa
Manitoba agriculture (2024) Soil Fertility Guide. https://www.gov.mb.ca/agriculture/crops/soil-fertility/soil-fertility-guide/general-fertilizer-recs-without-soil-test.html. Access 2024. 05. 15.


```{r adding data of three crops}
# Add a column to Sorghum
sorghum_rows <- N2O_long[N2O_long$Crop == "Corn", ]
sorghum_rows$Crop <- "Sorghum"

# Add a column to camelina 
camelina_rows <- N2O_long[N2O_long$Crop == "Canola", ]
camelina_rows$Crop <- "Camelina"

# Add a colum to triticale
triticale_rows <- N2O_long[N2O_long$Crop == "Spring wheat", ]
triticale_rows$Crop <- "Triticale"

N2O_long <- rbind(N2O_long, sorghum_rows, camelina_rows, triticale_rows)
N2O_long <- N2O_long[order(N2O_long$Region),]
#Save the output for CSED
#write.csv(N2O_long, "../Output/Direct N2O_CRN_Canada.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean = 0.006674, median = 0.005246(not fertilzer weighted)

#write.csv(N2O_long, "../Output/Direct N2O_CRN_Canada_WO Silage.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean = 0.006662, median = 0.005164(not fertilzer weighted)

```




```{r pivot to wide and export}
#Replace 0 to NA
N2O_long$N2O.IEF[N2O_long$N2O.IEF == 0] <- NA

# Pivot N2O.IEF
N2O_wide <- N2O_long %>%
  pivot_wider(names_from =  Region, values_from = N2O.IEF) #kg N2O-N per kg N

#Save the output for CSED
#write.csv(N2O_wide, "../Output/Direct N2O_FSN_Canada_wide.csv")


```


## Comparison 

1. Our CRF shows IEF for direct emission (crop and grassland) is about 0.006732489 kg N2O-N / kg 
 (Table 3.D from CAN_2023_2021)
 
2. In Open LCA model, they have EF1, the average EF1 is 0.009995  (3. N2O Emission Factors)
  

```{r result}

tibble(N2O_wide) 


#compare the original OpenLCA results
load("../Input/OpenLCA.RData") #OpenLCA

#Inner join OpenLCA to N2O_long for comarios
Direct_comp <- N2O_long %>%
  inner_join(OpenLCA, by = c ("Region" = "Region")) %>%
  mutate(Diff = ifelse(N2O.IEF != 0, N2O.IEF- EF1, 0)) %>%
  mutate(Diff.p = round(Diff/EF1*100,1))

#select data for pivot to wide, need to take p first, otherwise data was elminated from Leaching_comp
Direct_comp.p <- Direct_comp %>%
  select(Region, Crop, Diff.p)
Direct_comp <- Direct_comp %>%
  select(Region, Crop, Diff)


# Pivot Difference
Direct_comp_wide <- Direct_comp %>%
  pivot_wider(names_from =  Region, values_from = Diff) #kg N2O-N per kg N

Direct_comp.p_wide <- Direct_comp.p %>%
  pivot_wider(names_from =  Region, values_from = Diff.p) #kg N2O-N per kg N


```