---
title: "Direct N2O emissions from regionalized crop"
author: "Dr. Chih-Yu Hung"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Data descrpition
To calculate direct N2O emissions for regionalized crop, we need
1. Precipitation and Potential evaporation (INPUT-EF_Fixed_Para)
2. Topography (INPUT-EF_Fixed_Para)
3. Fraction of soil texture (INPUT-EF_dep_Para_texture)


## Calculate the BaseEF and EF

```{r EF caluclation}
source("Function for N2O EF_Direct.R")
load("../Input/CA_Direct_2016-2022.RData")
#The last 5 years
SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))

#Calculate the EF
Result <- N2OEF_direct(SiteData) 

#Maritime provinces
MT <- c("NL","NS","PE","NB")

N2O_long<- Result[[1]] %>%
  mutate (Region = case_when(         #combine the region
          Province %in% MT ~ "MT",
          TRUE ~ Province)) %>%
  group_by(Region,Crop) %>%
  summarise(N2O.IEF =  mean(N2O.IEF))

# Pivot N2O.IEF
N2O_wide <- N2O_long %>%
  pivot_wider(names_from =  Region, values_from = N2O.IEF) #kg N2O-N per kg N


#Save the output for CSED
write.csv(N2O_wide, "../Output/Direct N2O_FSN_Canada.csv")

```


## Comparison 

1. Our CRF shows IEF for volatilization is about 0.00739324 kg N2O-N / kg 
 (Table 3.D from CAN_2023_2021)
 
2. In Open LCA model, they have FracGASF and EF4, the average FracGASF is 0.068 and average EF4 is 0.010
  This means the average EF could be 0.00068. (3. N2O Emission Factors)
  
3. In my calculation, the average IEF is 0.0004357. This is lower than the two values above. 

```{r result}

tibble(N2O_wide, caption = "Average N2O IEF from volatilization for crops in 2018-2022") 


#compare the original OpenLCA results
load("../Input/OpenLCA.RData") #OpebLCA

#Inner join OpenLCA to N2O_long for comarios
Volatilization_comp <- N2O_long %>%
  inner_join(OpenLCA, by = c ("Region" = "Region")) %>%
  mutate(Diff = ifelse(N2O.IEF != 0, N2O.IEF- (FracLEACH*EF5), 0)) %>%
  mutate(Diff.p = round(Diff/(FracLEACH*EF5)*100,1))

#select data for pivot to wide, need to take p first, otherwise data was elminated from Leaching_comp
Volatilization_comp.p <- Volatilization_comp %>%
  select(Region, CropID, Diff.p)
Volatilization_comp <- Volatilization_comp %>%
  select(Region, CropID, Diff)


# Pivot Difference
Volatilization_comp_wide <- Volatilization_comp %>%
  pivot_wider(names_from =  Region, values_from = Diff) #kg N2O-N per kg N

Volatilization_comp.p_wide <- Volatilization_comp.p %>%
  pivot_wider(names_from =  Region, values_from = Diff.p) #kg N2O-N per kg N


```