---
title: "Direct N2O emissions from regionalized crop"
author: "Dr. Chih-Yu Hung"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
source("Function for N2O EF_Direct.R")
```

## Data descrpition
To calculate direct N2O emissions for regionalized crop, we need
1. Precipitation and Potential evaporation (INPUT-EF_Fixed_Para)
2. Topography (INPUT-EF_Fixed_Para)
3. Fraction of soil texture (INPUT-EF_dep_Para_texture)


## Add Sorghum, camelina and triticale data
In Canada, we have very little amount of grain sorghum. Considering the close relationship between grain Sorghum and corn and their similar N requirement and fertilizer management plan. We are going to use N2O EF of corn for sorghum. 

In Canada, the optimal fertilization plan for camelina was not determined. Recommended application range is 60-100 kg N /ha. However, CFIA (2017) suggested using fertilization plan of canola if further information is not available. 

Other province didn't have fertilizer guide for cameline (e.g. Manitoba Agriculture), AFOLU therefore used EFs of canola.  

Only QC, ON, SK, MB, AB (QC, ON and MB has very small amount <10000 metric ton) have triticale production


Ref:
Jordan B., Fertilizing to help maximize grain sorghum yields. https://www.cropscience.bayer.us/articles/channel/fertilizing-to-help-maximize-grain-sorghum-yields. access 24 May, 2024.
Vanderlip, R., Roozeboom, K., Fjell, D., Shroyer, J., Kok, H., Regehr, D., Whitney, D., Rogers, D.H., Alam, M., Jardine, D. Brooks, H.L., Taylor, R.K., Harner III, J.P., and Langemeier, L.N. 1998. Grain sorghum production handbook. C-687 (Revised). Kansas State University. https://bookstore.ksre.ksu.edu/pubs/c687.pdf.
Haag, L. (In cooperation with: Diaz, D. R., Obour, A., and Schlegel, A.) 2018. Grain sorghum fertility management. Kansas State University. https://www.northwest.k-state.edu/agronomy/documents/presentation_pdfs/2018%20Grain%20Sorghum%20Fertility%20Management%20SorghumU%202spp.pdf.
CFIA (2017) The Biology of Camelina sativa (L.) Crantz (Camelina). https://inspection.canada.ca/en/plant-varieties/plants-novel-traits/applicants/directive-94-08/biology-documents/camelina-sativa
Manitoba agriculture (2024) Soil Fertility Guide. https://www.gov.mb.ca/agriculture/crops/soil-fertility/soil-fertility-guide/general-fertilizer-recs-without-soil-test.html. Access 2024. 05. 15.


## GrainAgg (no silage and grain distinction)


```{r EF caluclation}

SiteData <-read_rds("../Input/CA_Direct_2016-2022_GrainAgg.rds") #Grain Agg


#The last 5 years
SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))

#Calculate the EF
Result <- N2OEF_direct(SiteData) 

#Atlantic provinces
AT <- c("NL","NS","PE","NB")

N2O_long<- Result[[1]] %>%
  mutate (Region = case_when(         #combine the region
          Province %in% AT ~ "AT",
          TRUE ~ Province)) %>%
  group_by(Region,Crop) %>%
  summarise(Tot.Fert = sum(Tot.Fert), Tot.N2O = sum(Tot.N2O)) %>%
  mutate(N2O.IEF =  round(ifelse(Tot.Fert >0, Tot.N2O/Tot.Fert,0),8))


# Add surrogate crops
N2O_long <- bind_rows(
  N2O_long,
  N2O_long %>% filter(Crop == "Corn") %>% mutate(Crop = "Sorghum"),
  N2O_long %>% filter(Crop == "Canola") %>% mutate(Crop = "Camelina"),
  N2O_long %>% filter(Crop == "Wheat(spring)") %>% mutate(Crop = "Triticale")
) %>%
  arrange(Region)


# Output summary and CSV
summary(N2O_long$N2O.IEF)
#mean = 0.007805, mediam = 0.006275(not fertilzer weighted)
write.csv(N2O_long, "../Output/Direct N2O_FSN_Canada_GrainAgg.csv", row.names = FALSE)
```




```{r adding data of three crops}

# Read Grain-split data
SiteData <- read_rds("../Input/CA_Direct_2016-2022_GrainSplit.rds")


# Use last 5 years
SiteData <- SiteData %>% filter(Year > (max(Year) - 5))

# Calculate EF
Result <- N2OEF_direct(SiteData)

# Group and summarize results
AT <- c("NL", "NS", "PE", "NB")
N2O_long <- Result[[1]] %>%
  mutate(Region = case_when(Province %in% AT ~ "AT", TRUE ~ Province)) %>%
  group_by(Region, Crop) %>%
  summarise(Tot.Fert = sum(Tot.Fert), Tot.N2O = sum(Tot.N2O), .groups = "drop") %>%
  mutate(N2O.IEF = round(ifelse(Tot.Fert > 0, Tot.N2O / Tot.Fert, 0), 8))

# Add surrogate crops
N2O_long <- bind_rows(
  N2O_long,
  N2O_long %>% filter(Crop == "Corn(silage)") %>% mutate(Crop = "Sorghum(silage)"),
  N2O_long %>% filter(Crop == "Corn(grain)") %>% mutate(Crop = "Sorghum(grain)"),
  N2O_long %>% filter(Crop == "Canola") %>% mutate(Crop = "Camelina"),
  N2O_long %>% filter(Crop == "Wheat(spring)") %>% mutate(Crop = "Triticale")
) %>%
  arrange(Region)

# Output summary and CSV
summary(N2O_long$N2O.IEF) #mean = 0.007872, mediam = 0.006220(not fertilzer weighted)
write.csv(N2O_long, "../Output/Direct N2O_FSN_Canada_GrainSplit.csv", row.names = FALSE)



```