---
title: "Direct N2O emissions from regionalized crop"
author: "Dr. Chih-Yu Hung"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Data descrpition
To calculate direct N2O emissions for regionalized crop, we need
1. Precipitation and Potential evaporation (INPUT-EF_Fixed_Para)
2. Topography (INPUT-EF_Fixed_Para)
3. Fraction of soil texture (INPUT-EF_dep_Para_texture)


```{r Data input from our database}
INPUT_Fixed <-read.csv("../Input/INPUT-EF_Fixed_Para.csv")
INPUT_texture <-read.csv("../Input/INPUT-EF_dep_para_texture.csv")

```

## Calculate the Base EF
Base EF is calculated based on the P and PE, fraction of topography and soil texture. The spatial unit here is ecodistrict.  
```{r BaseEF, echo=FALSE}
#Calcuate the weighted ratio factor for soil texutre
INPUT_texture <- INPUT_texture %>%
          mutate(Wtd_RF_TX = ((Frac_Coarse*0.49 + Frac_Medium*1 + Frac_Fine*2.55)/
                             (Frac_Coarse + Frac_Medium + Frac_Fine)))

#Base EF
INPUT_Fixed <- INPUT_Fixed %>%
  mutate(EF_Base_before_texture = ifelse(P>PE,exp(0.00558*P-7.7),
                                  exp(0.00558*PE-7.7)*F_TOPO+exp(0.00558*P-7.7)*(1-F_TOPO)))
#Introduce the Weighted ratio factor to cacluate Base emission factor
Base_EF <- INPUT_texture %>%
  filter (Inv_Year_ID == 2021) %>%
  inner_join(INPUT_Fixed, by = c("ECODISTRIC"="Ecodistrict_ID" ,
                                 "PROVINCE" = "Province_ID")) %>%
  mutate(Base_EF = Wtd_RF_TX * EF_Base_before_texture)


```


## Calculate the EF for each Ecodistrict, unit N2O-N per kg N fertilizer
For this part, we need 
1. ratio factor for non-growing season which is 1/0.634
2. ratio factor for N source, which is 1 for synthetic fertilizer
3. ratio factor for cropping system, which is 1 for annual crop
```{r EF for ecodistrcit, echo=FALSE}
EF <- Base_EF %>%
  select(ECODISTRIC,PROVINCE,Base_EF) %>%
  mutate(EF = Base_EF * 1/0.634 *1 *1 )

```


## We need to know the N fertilizer for each crop and each ecodistrict

There are cropIII_ID, we only need specific type of crops
We can load a list from "INPUT_Crop-TypeIII.csv"

```{r fertilizer for ecodistrcit, echo=FALSE}
load("../Input/Crop type and real Fertilizer_2016-2022.RData")
```



## Apply the synthetic fertilizer amount to EF
```{r N2O emissions, ech0=FALSE}
N2Oe <- Fert_crop %>%
  inner_join(EF, by = c("Ecodistrict_ID" = "ECODISTRIC",
                         "Province_ID"= "PROVINCE")) %>%
  mutate(N2O_crop = EF * Crop_Real_N)

```

## Obtain the Total emissions in each crop and Province
```{r N2O emissions, ech0=FALSE}
N2O_crop_prov <- N2Oe %>%
  group_by(Province_ID,CropIII_ID) %>%
  summarise(N2O_tot=sum(N2O_crop))

load("../Input/Crop_production_prov.RData")

N2O_crop_EF <- N2O_crop_prov %>%
  inner_join(Production, by = c("Province_ID" = "Province",
                                "CropIII_ID" = "CropID")) %>%
  mutate(EF = ifelse(Production >0, N2O_tot/Production,0)) #convert production from metric ton to kg
  #the unit is N2O-N per kg production crop

```



## Combine the Maritime provinces
```{r Maritime province}
MT <- c("NL","NS","PE","NB")
N2O_crop_EF_region <- N2O_crop_EF %>%
  mutate (Region = case_when(
          Province_ID %in% MT ~ "MT",
          TRUE ~Province_ID)) %>%
  group_by(Region,CropIII_ID,Crop) %>%
  summarise(EF = mean(EF))

```