---
title: "Direct N2O from irrigation_US"
author: "Dr. Chih-Yu Hung"
date: "2025-05-22"
output: html_document
---

```{r setup, include=FALSE}
source("../setup.R")
source("Function for N2O EF_Irrigation_US.R")

#US irrigation fraction
INPUT_irri <- read.csv(us_outputs("1.4_Irrigation fraction.csv"))

```

## N2O emissions due to irrigation

This methodology firstly determine the proportion of irrigation area in a ecodistrict and then apply a ratio factor (EF_PE/EF_P, must > 1) and then deduct 1

No need to consider irrigation in winter, since there is no crop planting in non-growing season that is applicable to non-growing season coef, i.e. sum(USData$DOUBLE == "Y" & USData$NGS == "Y") = 0 


```{r GrainAgg}
load("../Input/US_Direct_2018-2022_GrainAgg.RData") #USData 


SiteData <- USData %>%
  left_join(INPUT_irri, by = c("STATE", "CropID")) %>%
  mutate(Frac_IRRI = ifelse(is.na(Frac_IRRI),ifelse(P>PE,0,1),Frac_IRRI)) %>%
  select(-COMMODITY_DESC, -SHORT_DESC)

#save(SiteData,file = "../Input/US_Irrigation_2018-2022_GrainAgg.RData")      


Result_Irri <- N2OEF_Irri(SiteData)


#write.csv(Result_Irri$Result_Prov_Crop,"../Output/Direct N2O_Irrigation_US_GrainAgg.csv",row.names = FALSE)
summary(Result_Irri$Result_Prov_Crop$N2O.IEF)
#mean IEF = 0.0088761, Median = 0.0022571```

```


```{r GrainSplit}

load("../Input/US_Direct_2018-2022_GrainSplit.RData") #USData 


SiteData <- USData %>%
  left_join(INPUT_irri, by = c("STATE", "CropID")) %>%
  mutate(Frac_IRRI = ifelse(is.na(Frac_IRRI),ifelse(P>PE,0,1),Frac_IRRI)) %>%
  select(-COMMODITY_DESC, -SHORT_DESC)

#save(SiteData,file = "../Input/US_Irrigation_2018-2022_GrainSplit.RData")      


Result_Irri <- N2OEF_Irri(SiteData)


write.csv(Result_Irri$Result_Prov_Crop,"../Output/Direct N2O_Irrigation_US_GrainSplit.csv",row.names = FALSE)
ummary(Result_Irri$Result_Prov_Crop$N2O.IEF)
#mean IEF = 0.0103611, Median = 0.0029001```

```