---
title: "SiteData_prep_direct_CropResidue_US"
author: "Dr. Chih-Yu Hung"
date: "2025-07-10"
output: html_document
---

```{r setup, include=FALSE}
source("../setup.R")
USData_CRN <- read_csv(us_outputs("4_US_crop N input.csv")) %>%
              select(STATE, COUNTY, YEAR, CROP, CLASS, UTIL_PRACTICE_DESC, DOUBLE, CRN)

#Take the FSN data to have environmental factors
USData_FSN <- read_csv(us_outputs("0.1_US_Direct.csv"))


RawUSData <- USData_CRN %>%
  left_join(USData_FSN, by = c("STATE", "COUNTY", "YEAR", "CROP", "CLASS", "DOUBLE")) %>%
  mutate(Fert_LCA_kg = CRN)



```

## Objective

"CA_Direct_2016-2022.RData" has prepared necessary information to calculate N2O emissions from synthetic fertilizer. Now, we need to replace synthetic fertilizer N with crop residue N and change fertilizer factors (N sources)

### Example for SiteData
    RegionID:        suggest using HUP
    ProvinceID:      Province or State ID
    Year:            Inventory year
    CropID:          Name of crop
    Precipitation:   Precipitation (mm)
    Topo:            percentage of area in low topo
    Evapo:           Potential evapotranspiration (mm)
    Crop_coef:       characteristic,1 Annual, 2 Perennial, 3 impast, 4 unimpast 
    NSE:             1 or 0, 1 means non-growing season true
    NS:              Nitrogen source 1 inorganic, 2 animal manure, 3 Crop residue, 4 biosolid 
    Frac_C:          Fraction of soil texture in coarse texture
    Frac_M:          Fraction of soil texture in medium texture
    Frac_F:          Fraction of soil texture in fine texture
    Fertilizer_Applied 
    
```{r Data preparation}
for (silage_drop in c(FALSE, TRUE)) {

USData <- RawUSData %>% 
    { if (silage_drop) filter(.,UTIL_PRACTICE_DESC != "SILAGE") else . } 

# Apply the mapping to the dataset
USData <- USData %>%
  mutate(CropID = crop_mapping(CROP, CLASS)) %>%
  filter(!is.na(CropID))%>%  # Remove unwanted crops
  select(-CROP, -CLASS) %>%
  mutate(RegionID = str_c(CropID, STATE, COUNTY, YEAR, sep = "_"))%>% #add RegionID
  mutate(Frac_Coarse = Frac_Coarse/100) %>%
  mutate(Frac_Medium = Frac_Medium/100) %>%
  mutate(Frac_Fine = Frac_Fine/100)
  
#To correct NGS as if DOUBLE is true, no need to apply NGS coef. 
USData <- USData %>%
  mutate(NGS = if_else(DOUBLE == "Y" & NGS == "Y", "N", NGS))

#Crop coefficient, all annual in this project
USData$Crop_f <- 1

#Add NS data, all 2 crop residue in this project
USData$NSource <- 2

#replace NA in the fertilizer part if ever
USData[is.na(USData)] <- 0


#Save the sitedata
  saveRDS(
    USData,
    if (silage_drop)
      "../Input/US_Direct_Crop_2018-2022_WO_Silage.rds"
    else
      "../Input/US_Direct_Crop_2018-2022.rds"
  )

}
```





