---
title: "SiteData_prep_direct"
author: "Dr. Chih-Yu Hung"
date: "2024-05-17"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Objective

Preapre a `SiteData` file for `Direct N2O from synthetic fertilzer.Rmd`. 


### Example for SiteData
    RegionID:        suggest using HUP
    ProvinceID:      Province or State ID
    Year:            Inventory year
    CropID:          Name of crop
    Precipitation:   Precipitation (mm)
    Topo:            percentage of area in low topo
    Evapo:           Potential evapotranspiration (mm)
    Crop_coef:       characteristic,1 Annual, 2 Perennial, 3 impat, 4 unimpast 
    NSE:             1 or 0, 1 means non-growing season true
    NS:              Nitrogen source 1 inorganic, 2 animal manure, 3 Crop residue, 4 biosolid 
    Frac_C:          Fraction of soil texture in coarse texture
    Frac_M:          Fraction of soil texture in medium texture
    Frac_F:          Fraction of soil texture in fine texture
    
```{r Data preparation}


```