---
title: "SiteData_prep_direct"
author: "Dr. Chih-Yu Hung"
date: "2024-05-17"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Objective

Preapre a `SiteData` file for `Direct N2O from synthetic fertilzer.Rmd`. 


### Example for SiteData
    RegionID:        suggest using HUP
    ProvinceID:      Province or State ID
    Year:            Inventory year
    CropID:          Name of crop
    Precipitation:   Precipitation (mm)
    Topo:            percentage of area in low topo
    Evapo:           Potential evapotranspiration (mm)
    Crop_coef:       characteristic,1 Annual, 2 Perennial, 3 impast, 4 unimpast 
    NSE:             1 or 0, 1 means non-growing season true
    NS:              Nitrogen source 1 inorganic, 2 animal manure, 3 Crop residue, 4 biosolid 
    Frac_C:          Fraction of soil texture in coarse texture
    Frac_M:          Fraction of soil texture in medium texture
    Frac_F:          Fraction of soil texture in fine texture
    Fertilizer_Applied 
    
```{r Data preparation}
#Crop and fertilizer data
load("../Input/Crop type and real fertilizer_2016-2022.RData") #Fert_crop

SiteData <- Fert_crop %>%
  select(RegionID, Province_ID, Ecodistrict = Ecodistrict_ID, Year = Inv_Year_ID, CropID, Fertilizer_Applied = Crop_Real_N)

#Crop coefficient, all annual in this project
SiteData$Crop_f <- 1

#Fix parameter, 1981-2010 climate normal and topography
INPUT_Fixed <-  read.csv("../Input/INPUT-EF_Fixed_Para.csv") %>%
  select(Ecodistrict_ID, Province_ID, Topo = F_TOPO, Precipitation = P, Evapo = PE)

SiteData <- SiteData %>%
  inner_join(INPUT_Fixed, by = c("Ecodistrict" = "Ecodistrict_ID", "Province_ID" = "Province_ID"))
  

#Soil texture data
INPUT_texture <-read.csv("../Input/INPUT-EF_dep_para_texture.csv") %>%
  filter(Inv_Year_ID >= 2016) %>%
  select(-ANCOARSE, -ANMED, -ANFINE, -PERCOARSE, -PERMED,-PERFINE)

SiteData <- SiteData %>%
  inner_join(INPUT_texture, by = c("Ecodistrict" = "ECODISTRIC", "Province_ID" = "PROVINCE", "Year" = "Inv_Year_ID"))

#Add NSE data, all 1 in this project
SiteData$NSE <- 1
#Add NS data, all 1(synthetic fertilizer) in this project
SiteData$NSource <- 1

#Save the SiteData
save(SiteData, file = "../Input/CA_Direct_2016-2022.RData")
```





