---
title: "SiteData_prep_direct_CropResidue"
author: "Dr. Chih-Yu Hung"
date: "2025-06-25"
output: html_document
---

```{r setup, include=FALSE}
source("../setup.R")
RawsiteData <- read_xlsx("../Input/CR_Residual_N_output.xlsx") %>%
  select(Inv_Year_ID, Province_ID, CropIII_ID, Crop_TypeIII_Name, Ecodistrict_ID,Res_N)%>% #downsize the data
  filter(Inv_Year_ID >= 2016 & Inv_Year_ID <=2022) %>%
  filter(CropIII_ID %in% target_cropID) %>%
  mutate(RegionID = str_c(Crop_TypeIII_Name, Province_ID, Ecodistrict_ID, Inv_Year_ID, sep = "_")) %>%
  rename(Fertilizer_Applied = Res_N, Year = Inv_Year_ID, CropID = Crop_TypeIII_Name, Ecodistrict = Ecodistrict_ID)


```

## Objective

"CA_Direct_2016-2022.RData" has prepared necessary information to calculate N2O emissions from synthetic fertilizer. Now, we need to replace synthetic fertilizer N with crop residue N and change fertilizer factors (N sources)

### Example for SiteData
    RegionID:        suggest using HUP
    ProvinceID:      Province or State ID
    Year:            Inventory year
    CropID:          Name of crop
    Precipitation:   Precipitation (mm)
    Topo:            percentage of area in low topo
    Evapo:           Potential evapotranspiration (mm)
    Crop_coef:       characteristic,1 Annual, 2 Perennial, 3 impast, 4 unimpast 
    NSE:             1 or 0, 1 means non-growing season true
    NS:              Nitrogen source 1 inorganic, 2 animal manure, 3 Crop residue, 4 biosolid 
    Frac_C:          Fraction of soil texture in coarse texture
    Frac_M:          Fraction of soil texture in medium texture
    Frac_F:          Fraction of soil texture in fine texture
    Fertilizer_Applied 
    
```{r Data preparation}
for (silage_drop in c(FALSE, TRUE)) {

SiteData <- RawsiteData %>% 
    { if (silage_drop) filter(., CropIII_ID != 22) else . } 
#Crop coefficient, all annual in this project
SiteData$Crop_f <- 1

#Fix parameter, 1981-2010 climate normal and topography
INPUT_Fixed <-  read.csv("../Input/INPUT-EF_Fixed_Para.csv") %>%
  select(Ecodistrict_ID, Province_ID, Topo = F_TOPO, Precipitation = P, Evapo = PE)

SiteData <- SiteData %>%
  inner_join(INPUT_Fixed, by = c("Ecodistrict" = "Ecodistrict_ID", "Province_ID" = "Province_ID"))
  

#Soil texture data
INPUT_texture <-read.csv("../Input/4-3_EF_time_dep_para.csv") %>% #Data from 2024 NIR
    filter(Inv_Year_ID >= 2016) %>%
  select(Inv_Year_ID, ECODISTRIC, PROVINCE, A_PCoarse, A_PMed, A_PFine)  %>%
  rename( Frac_Coarse = `A_PCoarse`, Frac_Medium = `A_PMed`, Frac_Fine = `A_PFine`)



SiteData <- SiteData %>%
  inner_join(INPUT_texture, by = c("Ecodistrict" = "ECODISTRIC", "Province_ID" = "PROVINCE", "Year" = "Inv_Year_ID"))

#Add NSE data, all 1 in this project
SiteData$NSE <- 1
#Add NS data, all 2(crop residue) in this project
SiteData$NSource <- 2


#Merge silage and grain corn and fall and spring ryes
SiteData <- SiteData %>%
  # Harmonise the CropID
  mutate(
    CropID = str_replace_all(
      CropID,
     c(
        # any order of “corn” + “grain|silage”, case-insensitive
        "(?i)\\b(?:Corn\\s+(?:Grain|Silage)|(?:Grain|Silage)\\s+Corn)\\b" = "Corn",
        # any order of “rye”  + “spring|fall”,  case-insensitive
        "(?i)\\b(?:Rye\\s+(?:Spring|Fall)|(?:Spring|Fall)\\s+Rye)\\b"     = "Rye"
      )
    )
  ) %>% 
  group_by(Province_ID, Ecodistrict, Year, CropID) %>% 
  summarise(
    Fertilizer_Applied = sum(Fertilizer_Applied, na.rm = TRUE),
    across(-Fertilizer_Applied, first),
    .groups = "drop"
  ) %>% 
  mutate(
    RegionID = paste(CropID, Province_ID, Ecodistrict, Year, sep = "_")
  ) %>% 
  # put RegionID next to CropID 
  relocate(RegionID, .before = Province_ID)

#Save the sitedata
  saveRDS(
    SiteData,
    if (silage_drop)
      "../Input/CA_Direct_Crop_2016-2022_WO_Silage.rds"
    else
      "../Input/CA_Direct_Crop_2016-2022.rds"
  )

}
```





