---
title: "SiteData_prep_direct_US"
author: "Dr. Chih-Yu Hung"
date: "2025-03-02"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
USData <- read_csv("C:/Users/HungC/OneDrive - EC-EC/2023/Fuel LCA/US data/Outputs/0.1_US_Direct.csv")

```

## Objective

Preapre a `SiteData` file for `US_Direct N2O from synthetic fertilzer.Rmd`. 


### Example for SiteData
    RegionID:        suggest using HUP
    ProvinceID:      Province or State ID
    Year:            Inventory year
    CropID:          Name of crop
    Precipitation:   Precipitation (mm)
    Topo:            percentage of area in low topo
    Evapo:           Potential evapotranspiration (mm)
    Crop_coef:       characteristic,1 Annual, 2 Perennial, 3 impast, 4 unimpast 
    NSE:             1 or 0, 1 means non-growing season true
    NS:              Nitrogen source 1 inorganic, 2 animal manure, 3 Crop residue, 4 biosolid 
    Frac_C:          Fraction of soil texture in coarse texture
    Frac_M:          Fraction of soil texture in medium texture
    Frac_F:          Fraction of soil texture in fine texture
    Fertilizer_Applied 
    
```{r Data preparation}
#CropID needs to combine CROP and CLASS 
# Apply the mapping to the dataset
source("../Crop mapping.R")
USData <- USData %>%
  mutate(CropID = crop_mapping(CROP, CLASS)) %>%
  filter(!is.na(CropID))%>%  # Remove unwanted crops
  select(-CROP, -CLASS) %>%
  mutate(RegionID = str_c(CropID, STATE, COUNTY, YEAR, sep = "_"))%>% #add RegionID
  mutate(Frac_Coarse = Frac_Coarse/100) %>%
  mutate(Frac_Medium = Frac_Medium/100) %>%
  mutate(Frac_Fine = Frac_Fine/100)
  
#To correct NGS as if DOUBLE is true, no need to apply NGS coef. 
USData <- USData%>%
  mutate(NGS = if_else(DOUBLE == "Y" & NGS == "Y", "N", NGS))

#Crop coefficient, all annual in this project
USData$Crop_f <- 1

#Add NS data, all 1(synthetic fertilizer) in this project
USData$NSource <- 1

#replace NA in the fertilizer part
USData[is.na(USData)] <- 0

#Save the SiteData
save(USData, file = "../Input/US_Direct_2018-2022.RData")
```





