---
title: "SiteData_prep_direct_CropResidue_US"
author: "Dr. Chih-Yu Hung"
date: "2025-07-10"
output: html_document
---

```{r setup, include=FALSE}
source("../setup.R")
USData_CRN <- read_csv(us_outputs("4_US_crop N input.csv")) %>%
              select(STATE, COUNTY, YEAR, CROP, CLASS, UTIL_PRACTICE_DESC, DOUBLE, CRN)

#Take the FSN data to have environmental factors
USData_FSN <- read_csv(us_outputs("0.1_US_Direct.csv"))


RawUSData <- USData_CRN %>%
  left_join(USData_FSN, by = c("STATE", "COUNTY", "YEAR", "CROP", "CLASS", "DOUBLE")) %>%
  mutate(Fert_LCA_kg = CRN)
```

## Objective

"CA_Direct_2016-2022.RData" has prepared necessary information to calculate N2O emissions from synthetic fertilizer. Now, we need to replace synthetic fertilizer N with crop residue N and change fertilizer factors (N sources)

### Example for SiteData
    RegionID:        suggest using HUP
    ProvinceID:      Province or State ID
    Year:            Inventory year
    CropID:          Name of crop
    Precipitation:   Precipitation (mm)
    Topo:            percentage of area in low topo
    Evapo:           Potential evapotranspiration (mm)
    Crop_coef:       characteristic,1 Annual, 2 Perennial, 3 impast, 4 unimpast 
    NSE:             1 or 0, 1 means non-growing season true
    NS:              Nitrogen source 1 inorganic, 2 animal manure, 3 Crop residue, 4 biosolid 
    Frac_C:          Fraction of soil texture in coarse texture
    Frac_M:          Fraction of soil texture in medium texture
    Frac_F:          Fraction of soil texture in fine texture
    Fertilizer_Applied 
    
```{r Data preparation}
for (GrainAgg in c(FALSE, TRUE)) {

USData <- RawUSData

# Apply the mapping to the dataset
USData <- USData %>%
  mutate(CropID = crop_mapping(CROP, CLASS, UTIL_PRACTICE_DESC)) %>%
  filter(!is.na(CropID))%>%  # Remove unwanted crops
  select(-CLASS) %>%
  mutate(Frac_Coarse = Frac_Coarse/100) %>%
  mutate(Frac_Medium = Frac_Medium/100) %>%
  mutate(Frac_Fine = Frac_Fine/100)
  
#To correct NGS as if DOUBLE is true, no need to apply NGS coef. 
USData <- USData %>%
  mutate(NGS = if_else(DOUBLE == "Y" & NGS == "Y", "N", NGS))

#Crop coefficient, all annual in this project
USData$Crop_f <- 1

#Add NS data, all 2 crop residue in this project
USData$NSource <- 2

#replace NA in the fertilizer part if ever
USData[is.na(USData)] <- 0


#Aggregate
  if (GrainAgg) {
    USData <- USData %>%
      mutate(CropID = str_remove(CropID, "\\(grain\\)|\\(silage\\)")) %>%
      # Generalize crop name by removing silage/grain in parentheses
      group_by(STATE, COUNTY, YEAR, DOUBLE, CropID) %>%
      summarise(
        Fert_LCA_kg = sum(Fert_LCA_kg, na.rm = TRUE),
        PE  = first(PE),
        P   = first(P),
        Topo = first(Topo),
        NGS = first(NGS),
        Frac_Fine = first(Frac_Fine),
        Frac_Medium = first(Frac_Medium),
        Frac_Coarse = first(Frac_Coarse),
        Crop_f = first(Crop_f),
        NSource = first(NSource),
        .groups = "drop"
      ) %>%
      mutate(RegionID = str_c(CropID, STATE, COUNTY, YEAR, sep = "_")) %>%
      relocate(RegionID, .before = STATE)
  } else {
    # When NOT aggregating, preserve individual crop types
    USData <- USData %>%
      mutate(RegionID = str_c(CropID, STATE, COUNTY, YEAR, sep = "_")) %>%
      relocate(RegionID, .before = STATE)
  }



#Save the sitedata
  saveRDS(
    USData,
    if (GrainAgg)
      "../Input/US_Direct_Crop_2018-2022_GrainAgg.rds"
    else
      "../Input/US_Direct_Crop_2018-2022_GrainSplit.rds"
  )

}
```





