---
title: "Ammonia Volatilization Emission Factor"
author: "Chih-Yu Hung"
date: "2024-02-27, Revised in May"
output: html_document
knit: (function(input, encoding) {
    rmarkdown::render(
      input = input,
      encoding = encoding,
      envir = globalenv()
    )
  })
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse);library(xlsx);library(openxlsx)
source("Function for Indirect N2O EF_Volatilization.R")
load("../Input/CA_volatilization_2016-2022.RData") #SiteData
SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))
# Data preparation
CA_volatilization <- Volatilization_prep(SiteData) #2018-2022

```

## Objectives
Estimate indirect N2O emissions from volatilization.
Input data:
Real fertilizer usage: Input/Crop type and real Fertilizer_2016-2022.RData
Function:
Function for Indirect N2O EF_volatilization


``` {r EF wide table}

Result_vola <- N2OEF_volatilization(CA_volatilization)

#save excel file
# for (i in 1:length(Result_vola)){
#  write.xlsx(Result_vola[[i]],
#              file = "Output/CA_volatilization.xlsx",sheetName = names(Result_vola[i]),row.names = F)
#   }


# Pivot NH3.IEF
NH3_pivot <-Result_vola[[1]] %>%
  select(ProvinceID, CropID, NH3.IEF) %>%
  pivot_wider(names_from = ProvinceID, values_from = NH3.IEF)

# Pivot N2O.IEF
N2O_pivot <- Result_vola[[1]] %>%
  select(ProvinceID, CropID, N2O.IEF) %>%
  pivot_wider(names_from =  ProvinceID, values_from = N2O.IEF)


```

## For the Clean Fuel Regulation

For the regulation and LCA model, AFOLU needs to provide a table of volatilized N2O EF, which has EFs for provinceas (Maritime) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Maritime provinces. 

```{r CFR and LCA model}
#Maritime provinces
MT <- c("NL","NS","PE","NB")

N2O_long<- Result_vola[[4]] %>%
  mutate (Region = case_when(         #combine the region
          ProvinceID %in% MT ~ "MT",
          TRUE ~ProvinceID)) %>%
  group_by(Region,CropID) %>%
  summarise(N2O.IEF = ifelse(sum(Fertilizer_Applied) > 0, sum(N2O)/sum(Fertilizer_Applied), 0))


```
## Add Sorghum data
In Canada, we have very little amount of grain sorghum. Considering the close relationship between grain Sorghum and corn and their similar N requirement and fertilizer management plan. We are going to use N2O EF of corn for sorghum. 

Ref:
Jordan B., Fertilizing to help maximize grain sorghum yields. https://www.cropscience.bayer.us/articles/channel/fertilizing-to-help-maximize-grain-sorghum-yields. access 24 May, 2024.
Vanderlip, R., Roozeboom, K., Fjell, D., Shroyer, J., Kok, H., Regehr, D., Whitney, D., Rogers, D.H., Alam, M., Jardine, D. Brooks, H.L., Taylor, R.K., Harner III, J.P., and Langemeier, L.N. 1998. Grain sorghum production handbook. C-687 (Revised). Kansas State University. https://bookstore.ksre.ksu.edu/pubs/c687.pdf.
Haag, L. (In cooperation with: Diaz, D. R., Obour, A., and Schlegel, A.) 2018. Grain sorghum fertility management. Kansas State University. https://www.northwest.k-state.edu/agronomy/documents/presentation_pdfs/2018%20Grain%20Sorghum%20Fertility%20Management%20SorghumU%202spp.pdf.

```{r adding sorghum data}
corn_rows <- N2O_long[N2O_long$CropID == "Corn(grain)", ]
# Modify the Crop column to Sorghum
sorghum_rows <- corn_rows
sorghum_rows$CropID <- "Sorghum"
N2O_long <- rbind(N2O_long, sorghum_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

```

## Add Camelina data
In Canada, the optimal fertilization plan for camelina was not determined. Recommended application range is 60-100 kg N /ha. However, CFIA (2017) suggested using fertilization plan of canola if further information is not available. 

SK's fertilizer guide showed camelina is planted in brown soil and their default fertilizer rate is 75 kg N /ha in brown soils, which is lower than 85 kg N /ha for canola in brown soils in SK. Therefore we applied a correction factor 75/85 to the N2O EF in SK. (Agriculture Saskatchewan, 2024)

Other province didn't have fertilizer guide for cameline (e.g. Manitoba Agriculture), AFOLU therefore used EFs of canola.  


CFIA (2017) The Biology of Camelina sativa (L.) Crantz (Camelina). https://inspection.canada.ca/en/plant-varieties/plants-novel-traits/applicants/directive-94-08/biology-documents/camelina-sativa
Manitoba agriculture (2024) Soil Fertility Guide. https://www.gov.mb.ca/agriculture/crops/soil-fertility/soil-fertility-guide/general-fertilizer-recs-without-soil-test.html. Access 2024. 05. 15.
Agriculture Saskatchewan (2024) Crop planning guide online calculator 2024. 
https://publications.saskatchewan.ca/#/products/122662

```{r adding Camelina data}
canola_rows <- N2O_long[N2O_long$CropID == "Canola", ]
# Modify the Crop column to Sorghum
camelina_rows <- canola_rows
camelina_rows$CropID <- "Camelina"
#camelina_rows$N2O.IEF[camelina_rows$Region == "AB"] <- camelina_rows$N2O.IEF[camelina_rows$Region == "AB"] *0.9
camelina_rows$N2O.IEF[camelina_rows$Region == "SK"] <- camelina_rows$N2O.IEF[camelina_rows$Region == "SK"] * (75/85)

N2O_long <- rbind(N2O_long, camelina_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

```


## Add Triticale data
Only QC, ON, SK, MB, AB (QC, ON and MB has very small amount <10000 metric ton) have triticale production

Government of Alberta - Ministry of Agriculture and Irrigation (2004) recommends to use 35-90 (lb/acre) for triticale and 35-110 (lb/acre) for spring wheat on irrigated lands. AFOLU assumed avg. fertilization rate for triticale in AB is 62.5 lb/acre, which is 86% of fertilization rate of spring wheat. Since there is no documents to show the actual fertilizer usage on triticale production, AFOLU applied 86% as correction factor for triticale. 


Manitoba Agriculture (2024) recommends to use less fertilizer for triticale than wheat. The reduction rate from 0.67 to 0.73 depends on the previous crop, (e.g. fallow and stubble 0.67, and after grass 0.73). There is no information to know the previous crops for triticale field. A adjustment factor 0.7 is applied to triticale fertilization in Manitoba. 



Government of Alberta - Ministry of Agriculture and Irrigation (2004) Alberta Fertilizer Guide. https://open.alberta.ca/publications/2909624   
Agriculture Manitoba (2024) Triticale Production and Management
https://www.gov.mb.ca/agriculture/crops/crop-management/triticale.html access on June 10, 2024
Agriculture Manitoba (2024) Spring Wheat Production and Management. 
https://www.gov.mb.ca/agriculture/crops/crop-management/spring-wheat.html access on June 10, 2024

```{r adding Tritical data}
wheat_rows <- N2O_long[N2O_long$CropID == "Wheat(spring)", ]
# Modify the Crop column to Sorghum
triticale_rows <- wheat_rows
triticale_rows$CropID <- "Triticale"
triticale_rows$N2O.IEF[triticale_rows$Region == "AB"] <- triticale_rows$N2O.IEF[triticale_rows$Region == "AB"] *0.86
triticale_rows$N2O.IEF[triticale_rows$Region == "MB"] <- triticale_rows$N2O.IEF[triticale_rows$Region == "MB"] * 0.7

N2O_long <- rbind(N2O_long, triticale_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

```

```{r pivot to wide and export table}
#Replace 0 to NA
N2O_long$N2O.IEF[N2O_long$N2O.IEF == 0] <- NA

# Pivot N2O.IEF
N2O_wide <- N2O_long %>%
  pivot_wider(names_from =  Region, values_from = N2O.IEF) #kg N2O-N per kg N


#Save the output for CSED
write.csv(N2O_wide, "../Output/Indirect N2O_volatilization_Canada.csv")


```


## Comparison 

1. Our CRF shows IEF for volatilization is about 0.00739324 kg N2O-N / kg 
 (Table 3.D from CAN_2023_2021)
 
2. In Open LCA model, they have FracGASF and EF4, the average FracGASF is 0.068 and average EF4 is 0.010
  This means the average EF could be 0.00068. (3. N2O Emission Factors)
  
3. In my calculation, the average IEF is 0.0004357. This is lower than the two values above. 

```{r result}

tibble(N2O_wide, caption = "Average N2O IEF from volatilization for crops in 2018-2022") 


#compare the original OpenLCA results
load("../Input/OpenLCA.RData") #OpebLCA

#Inner join OpenLCA to N2O_long for comarios
Volatilization_comp <- N2O_long %>%
  inner_join(OpenLCA, by = c ("Region" = "Region")) %>%
  mutate(Diff = ifelse(N2O.IEF != 0, N2O.IEF- (FracLEACH*EF5), 0)) %>%
  mutate(Diff.p = round(Diff/(FracLEACH*EF5)*100,1))

#select data for pivot to wide, need to take p first, otherwise data was elminated from Leaching_comp
Volatilization_comp.p <- Volatilization_comp %>%
  select(Region, CropID, Diff.p)
Volatilization_comp <- Volatilization_comp %>%
  select(Region, CropID, Diff)


# Pivot Difference
Volatilization_comp_wide <- Volatilization_comp %>%
  pivot_wider(names_from =  Region, values_from = Diff) #kg N2O-N per kg N

Volatilization_comp.p_wide <- Volatilization_comp.p %>%
  pivot_wider(names_from =  Region, values_from = Diff.p) #kg N2O-N per kg N


```


