---
title: "Ammonia Volatilization Emission Factor"
author: "Chih-Yu Hung"
date: "2024-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Objectives
Estimate indirect N2O emissions (volatilization)
Sum of relevant coefficient for each ecodistricts
The goal is to calculate sum of relevant coefficients in every ecodistrcit
The weighted value in ecodistrict to province will be based on fertilizer amount 

```{r Read data and explanation}
Soil <- read.csv("../Input/ACTIVITYDATA-EcoDistrict_Soils.csv") #Ecod,pH,CEC,proportion
Fert <- read.csv("../Input/ACTIVITYDATA-Fertilzer_Proportion.csv") #Year, Prov, Fert, proportion
EF4 <- read.csv("../Input/EF4.csv") #Ecod, Ecozone, Prov
App <- read.csv("../Input/Application method.csv") #Year, Prov, Ecod, application, App amount
Fert.amount <- read.csv("../Input/N2OCOMP_Indirect_Emissions_ED_with_Descr.csv")
App.prop <- read.csv("../Input/Fertilizer_Propor_Methods.csv") #Fertilzer type, application type, coast and proportion

#Downsize the Fert.amount, We need FSN only.
Fert.amount <- Fert.amount[Fert.amount$N_SOURCE == "FSN" & Fert.amount$Crop_System == "Annual",]
Fert.amount <- Fert.amount[,c(1:3,10)]
```

## Coefficients types
1. Crop
2. Fertilizer: urea, UAN, Anhydrous ammonia, Other
3. Application: Broadcast and Incorporated
4. Soil pH: <7.25 and 7.25-8.5
5. Soil CEC: <250 mmol kg-1 and >250 mmol kg-1
6. Climate: temperature

```{r EF table}
#Add a column to App as 

#Merge data for fertilizers 
App_Fert <- App %>%
  inner_join(Fert, by = c("Year" = "Inv_Year", "Province_Sale_ID" = "Province_Sale_ID", "Fert_ID" = "Fert_ID")) 

#Merge data for soil conditions
App_Fert_Soil <- App_Fert %>%
  inner_join(Soil, by = c("Province_ID" = "Province", "EcoDistrict" = "EcoDistrict")) 


# Add application method proportion
West <- c("BC", "AB","SK","MB")

App_Fert_Soil <- App_Fert_Soil %>%
  mutate(Coast = case_when(
    Province_ID %in% West ~ "West",
    !(Province_ID %in% West) ~ "East"
  )
  )

App_Fert_Soil <- App_Fert_Soil %>%
  inner_join(App.prop, by = c("Coast" = "Coast", "Fert_ID" = "Fertilizer_ID", "Method_ID" = "Method_ID"))

#get rid of columns I don't need
App_Fert_Soil <- App_Fert_Soil %>%
  select(-Province_Sale_ID,-Fert_ID,-Crop_Type,-Emissions,-Coast,-ID)


#Add corrected applied fertilizer amount in the specific year and ecodistrict
App_Fert_Soil <- App_Fert_Soil %>%
  inner_join(Fert.amount, by = c("Year" = "Inv_Year_ID", "Province_ID" = "Province_ID", "EcoDistrict" = "Ecodistrict_ID")) %>%
  select(-Fertilizer_Applied, Fertilizer_Applied = N_Applied_kg)


```


## convert the parameters to coefficients
After the inner_join, there are three proportion/percentage/fraction in the table 
Fraction is the fraction of four types of fertilizer in the provinces 
Percentage is the percentage of soil condition in the Ecodistricts
Proportion is the proportion of three application methods in the provinces

### 1. Fertilizer types
There are four types of fertilizers. Convert them to coef.Urea(0.666), UAN(0.282),An ammonia(-1.151), Ohter(-0.238)

### 2. Convert application method
There are three methods, but 2. injection and 3. incorporated share common coef (-1.895)
The Broadcast has higher coef (-1.305)

### 3. Soil Condition
Soil condition can be pH < 7.25 (-1 convert to -1), >7.25 (0 convert to -0.608)
or CEC <250 (-1 convert to 0.0507) and CEC >250 (0 convert to 0.0848)

```{r intorduce parameters}
# 1. Fertilizer types
App_Fert_Soil <- App_Fert_Soil %>%
  mutate(Fert_coef = case_when(
    Fert_Name == "Urea" ~ 0.666,
    Fert_Name == "Urea ammonium nitrate" ~ 0.282,
    Fert_Name == "Anhydrous ammonia" ~ -1.151,
    Fert_Name == "Other" ~ -0.238,
    TRUE ~ 99
  ))

#Check error
if (any(is.na(App_Fert_Soil$Fert_coef))) {
    stop("Check Fertilizer results")
  } else {
  cat("All value converted")
  }
if(any(App_Fert_Soil$Fert_coef == 99)) {
  stop("Check Fertilizer results")
} else {
  cat("All value converted")
}


# 2. Convert application method
App_Fert_Soil <- App_Fert_Soil %>%
  mutate(Method_ID_coef = case_when(
    Method_ID == 1 ~ -1.305,
    Method_ID %in% c(2,3) ~ -1.895,
    TRUE ~ 99
  ))

#Check error
if (any(is.na(App_Fert_Soil$Method_ID_coef))) {
    stop("Check application results")
  } else {
  cat("All value converted")
  }
if(any(App_Fert_Soil$Method_ID_coef == 99)) {
  stop("Check application results")
} else {
  cat("All value converted")
}

# 3. Convert Soil condition
App_Fert_Soil <- App_Fert_Soil %>%
  mutate(SoilpH = case_when(
    Acid_Soil == -1 ~ -1,
    Acid_Soil == 0  ~ -0.608,
    TRUE ~ 99
  ))

App_Fert_Soil <- App_Fert_Soil %>%
  mutate(SoilCEC = case_when(
    Poor_Soil == -1 ~ 0.0507,
    Poor_Soil == 0  ~ 0.0848,
    TRUE ~ 99
  ))

#Check error
if (any(is.na(App_Fert_Soil$SoilpH)|is.na(App_Fert_Soil$SoilCEC))) {
    stop("Check soil results")
  } else {
  cat("All value converted")
  }
if(any(App_Fert_Soil$soilpH == 99|App_Fert_Soil$soilCEC == 99)) {
  stop("Check soil results")
} else {
  cat("All value converted")
}


#drop columns I don't need
Coef_vola <- App_Fert_Soil %>%
     select(-Method_ID, -Fert_Name, -Acid_Soil, -Poor_Soil)




```

##Prepare Ecodistrict and Ecozone data
``` {r Ecozone}
#Merge data for soil conditions
Coef_vola <- Coef_vola %>%
  inner_join(EF4, by = c("Province_ID" = "Province_ID", "EcoDistrict" = "ECODISTRIC")) 

#Assign EF4 for ecozones.
low.Ecozone <- c("Taiga Plain","Boreal Plains","Prairie","Montane Cordillera")
high.Ecozone <- c("Boreal Shield","Atlantic Maritime","MixedWood Plain","Pacific Maritime")
Coef_vola <- Coef_vola %>%
  mutate(EF4 = case_when(
    ZONE_NAME %in% low.Ecozone ~ 0.005,
    ZONE_NAME %in% high.Ecozone ~ 0.014,
    TRUE ~ 99
  )
  )

#Check error
if (any(is.na(Coef_vola$ECOZONE))) {
    stop("Check Ecozone results")
  } else {
  cat("All value converted")
  }
if(any(Coef_vola$EF4 == 99)) {
  stop("Check Ecozone results")
} else {
  cat("All value converted")
}

#drop unwanted columns
Coef_vola <- Coef_vola %>%
  select(-ECOZONE,-ZONE_NAME)

```



## Calculate the sum of coefficient

```{r Coefficient calculation}
#Calculate the sum of coefficient
#Sum of coefficient weighted by fraction of fertilizer, percentage of soil condition, and application method
Coef_vola_ecodist<- Coef_vola %>%
  mutate(Coef_sum =exp( -0.045 + Fert_coef + Method_ID_coef + SoilpH + SoilCEC -0.402) *Percentage*Fraction*Proportion) 


Coef_ecodist <- Coef_vola_ecodist %>%
  group_by(Year,Province_ID,EcoDistrict) %>%
  summarise(Avg.Coef = mean(Coef_sum),
         EF4 = mean(EF4),
         Fertilizer_Applied = mean(Fertilizer_Applied),
         Avg.NH3 = mean(Coef_sum*Fertilizer_Applied))

aa <- Coef_ecodist[Coef_ecodist$Year == 1992 & Coef_ecodist$Province_ID =="SK",]



Coef_province <-Coef_ecodist %>%
  group_by(Year, Province_ID) %>%
  summarise(Avg.NH3 = mean(Avg.Coef*Fertilizer_Applied),
            Avg.N2O = mean(Avg.Coef*Fertilizer_Applied*EF4),
            NH3.IEF = sum(Avg.Coef*Fertilizer_Applied)/sum(Fertilizer_Applied),
            N2O.IEF = sum(Avg.Coef*Fertilizer_Applied*EF4)/sum(Fertilizer_Applied))


  group_by(Year, Province_ID, Fertilizer_Applied,EcoDistrict, Fraction, Fert_coef, EF4) %>%
  summarise(Wd_Coef_sum = sum(Coef_sum*Percentage))

a <- Coef_vola_ecodist[Coef_vola_ecodist$Year == 1990 & Coef_vola_ecodist$Province_ID == "AB" & Coef_vola_ecodist$EcoDistrict == 622,]

aa <- Coef_vola %>%
  mutate(Coef_sum =exp( -0.045 + Fert_coef + Method_ID_coef + SoilpH + SoilCEC -0.402))

```