---
title: "Ammonia Volatilization Emission Factor"
author: "Chih-Yu Hung"
date: "2024-02-27, Revised in May"
output: html_document
knit: (function(input, encoding) {
    rmarkdown::render(
      input = input,
      encoding = encoding,
      envir = globalenv()
    )
  })
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)#library(xlsx)
source("Function for Indirect N2O EF_Volatilization.R")
load("../Input/CA_volatilization_2016-2022.RData") #SiteData
SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))
# Data preparation
CA_volatilization <- Volatilization_prep(SiteData) #2018-2022
#NEED to run "Function for Indirect N2O EF_volatilization"

```

## Objectives
Estimate indirect N2O emissions from volatilization.
Input data:
Real fertilizer usage: Input/Crop type and real Fertilizer_2016-2022.RData
Function:
Function for Indirect N2O EF_volatilization


``` {r EF wide table}

Result_vola <- N2OEF_volatilization(CA_volatilization)

#save excel file
#NEED JAVA waiting FOR IT REPSONSE, MAY 8, was told installed but failed May 14
# for (i in 1:3){
# write.xlsx(Result_vola[[i]],
#            file = "Output/CA_volatilization.xlsx",sheetName = names(Result_vola[i]),row.names = F, append =TRUE)
# }

# Pivot NH3.IEF
NH3_pivot <-Result_vola[[1]] %>%
  select(ProvinceID, CropID, NH3.IEF) %>%
  pivot_wider(names_from = ProvinceID, values_from = NH3.IEF)

# Pivot N2O.IEF
N2O_pivot <- Result_vola[[1]] %>%
  select(ProvinceID, CropID, N2O.IEF) %>%
  pivot_wider(names_from =  ProvinceID, values_from = N2O.IEF)


```

## For the Clean Fuel Regulation

For the regulation and LCA model, AFOLU needs to provide a table of volatilized N2O EF, which has EFs for provinceas (Maritime) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Maritime provinces. 

```{r CFR and LCA model}
#Maritime provinces
MT <- c("NL","NS","PE","NB")

N2O_long<- Result_vola[[4]] %>%
  mutate (Region = case_when(         #combine the region
          ProvinceID %in% MT ~ "MT",
          TRUE ~ProvinceID)) %>%
  group_by(Region,CropID) %>%
  summarise(N2O.IEF = ifelse(sum(Fertilizer_Applied) > 0, sum(N2O)/sum(Fertilizer_Applied), 0))

# Pivot N2O.IEF
N2O_wide <- N2O_long %>%
  pivot_wider(names_from =  Region, values_from = N2O.IEF) #kg N2O-N per kg N


#Save the output for CSED
write.csv(N2O_wide, "../Output/Indirect N2O_volatilization_Canada.csv")

```



## Comparison 

1. Our CRF shows IEF for volatilization is about 0.00739324 kg N2O-N / kg 
 (Table 3.D from CAN_2023_2021)
 
2. In Open LCA model, they have FracGASF and EF4, the average FracGASF is 0.068 and average EF4 is 0.010
  This means the average EF could be 0.00068. (3. N2O Emission Factors)
  
3. In my calculation, the average IEF is 0.0004357. This is lower than the two values above. 

```{r result}

tibble(N2O_wide, caption = "Average N2O IEF from volatilization for crops in 2018-2022") 


#compare the original OpenLCA results
load("../Input/OpenLCA.RData") #OpebLCA

#Inner join OpenLCA to N2O_long for comarios
Volatilization_comp <- N2O_long %>%
  inner_join(OpenLCA, by = c ("Region" = "Region")) %>%
  mutate(Diff = ifelse(N2O.IEF != 0, N2O.IEF- (FracLEACH*EF5), 0)) %>%
  mutate(Diff.p = round(Diff/(FracLEACH*EF5)*100,1))

#select data for pivot to wide, need to take p first, otherwise data was elminated from Leaching_comp
Volatilization_comp.p <- Volatilization_comp %>%
  select(Region, CropID, Diff.p)
Volatilization_comp <- Volatilization_comp %>%
  select(Region, CropID, Diff)


# Pivot Difference
Volatilization_comp_wide <- Volatilization_comp %>%
  pivot_wider(names_from =  Region, values_from = Diff) #kg N2O-N per kg N

Volatilization_comp.p_wide <- Volatilization_comp.p %>%
  pivot_wider(names_from =  Region, values_from = Diff.p) #kg N2O-N per kg N


```


