---
title: "Ammonia Volatilization Emission Factor"
author: "Chih-Yu Hung"
date: "2024-02-27, Revised in May"
output: html_document
knit: (function(input, encoding) {
    rmarkdown::render(
      input = input,
      encoding = encoding,
      envir = globalenv()
    )
  })
---

## Objectives
Estimate indirect N2O emissions from volatilization.
Input data:
Real fertilizer usage: Input/Crop type and real Fertilizer_2016-2022.RData
Function:
Function for Indirect N2O EF_volatilization


## Add Sorghum, Camelina, Triticale data
In Canada, we have very little amount of grain sorghum. Considering the close relationship between grain Sorghum and corn and their similar N requirement and fertilizer management plan. We are going to use N2O EF of corn for sorghum. 

In Canada, the optimal fertilization plan for camelina was not determined. Recommended application range is 60-100 kg N /ha. However, CFIA (2017) suggested using fertilization plan of canola if further information is not available. 

Other province didn't have fertilizer guide for cameline (e.g. Manitoba Agriculture), AFOLU therefore used EFs of canola.  



## Add Triticale data
Only QC, ON, SK, MB, AB (QC, ON and MB has very small amount <10000 metric ton) have triticale production


Ref:
Jordan B., Fertilizing to help maximize grain sorghum yields. https://www.cropscience.bayer.us/articles/channel/fertilizing-to-help-maximize-grain-sorghum-yields. access 24 May, 2024.
Vanderlip, R., Roozeboom, K., Fjell, D., Shroyer, J., Kok, H., Regehr, D., Whitney, D., Rogers, D.H., Alam, M., Jardine, D. Brooks, H.L., Taylor, R.K., Harner III, J.P., and Langemeier, L.N. 1998. Grain sorghum production handbook. C-687 (Revised). Kansas State University. https://bookstore.ksre.ksu.edu/pubs/c687.pdf.
Haag, L. (In cooperation with: Diaz, D. R., Obour, A., and Schlegel, A.) 2018. Grain sorghum fertility management. Kansas State University. https://www.northwest.k-state.edu/agronomy/documents/presentation_pdfs/2018%20Grain%20Sorghum%20Fertility%20Management%20SorghumU%202spp.pdf.
CFIA (2017) The Biology of Camelina sativa (L.) Crantz (Camelina). https://inspection.canada.ca/en/plant-varieties/plants-novel-traits/applicants/directive-94-08/biology-documents/camelina-sativa
Manitoba agriculture (2024) Soil Fertility Guide. https://www.gov.mb.ca/agriculture/crops/soil-fertility/soil-fertility-guide/general-fertilizer-recs-without-soil-test.html. Access 2024. 05. 15.
Agriculture Saskatchewan (2024) Crop planning guide online calculator 2024. 
https://publications.saskatchewan.ca/#/products/122662

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse);library(xlsx);library(openxlsx)
source("Function for Indirect N2O EF_Volatilization.R")

#Grain Agg
load("../Input/CA_volatilization_2016-2022_GrainAgg.RData") #SiteData
SiteData_GrainAgg <- SiteData_GrainAgg %>%
  filter (Year > (max(Year)-5))#2018-2022

#Grain Split
load("../Input/CA_volatilization_2016-2022_GrainSplit.RData") #SiteData
SiteData_GrainSplit <- SiteData_GrainSplit %>%
  filter (Year > (max(Year)-5))#2018-2022


# For the Clean Fuel Regulation
# For the regulation and LCA model, AFOLU needs to provide a table of Leached N2O EF, which has EFs for provinceas (Atlantic) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Atlantic provinces. 
AT <- c("NL","NS","PE","NB")

# Data preparation
CA_volatilization_GrainAgg <- Volatilization_prep(SiteData_GrainAgg) %>%
                              mutate (ProvinceID = case_when(         #combine the region
                              ProvinceID %in% AT ~ "AT",
                              TRUE ~ ProvinceID))

CA_volatilization_GrainSplit <- Volatilization_prep(SiteData_GrainSplit) %>%
                              mutate (ProvinceID = case_when(         #combine the region
                              ProvinceID %in% AT ~ "AT",
                              TRUE ~ ProvinceID))


```


``` {r EF wide table}

#GrainAgg
Result_vola_GrainAgg <- N2OEF_volatilization(CA_volatilization_GrainAgg)

N2O_long <- Result_vola_GrainAgg$Coef_Prov_Crop 

N2O_long <- bind_rows(
  N2O_long,
  N2O_long %>% filter(CropID == "Corn") %>% mutate(CropID = "Sorghum"),
  N2O_long %>% filter(CropID == "Canola") %>% mutate(CropID = "Camelina"),
  N2O_long %>% filter(CropID == "Wheat(spring)") %>% mutate(CropID = "Triticale")
) %>%
  arrange(ProvinceID)

#Save the output for CSED
#write.csv(N2O_long, "../Output/Indirect N2O_FSN_Volatilization_Canada_GrainAgg.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean 0.0004819, median, 0.0003313


#GrainSplit
Result_vola_GrainSplit <- N2OEF_volatilization(CA_volatilization_GrainSplit)
N2O_long <- Result_vola_GrainSplit$Coef_Prov_Crop 

# Add surrogate crops
N2O_long <- bind_rows(
  N2O_long,
  N2O_long %>% filter(CropID == "Corn(silage)") %>% mutate(CropID = "Sorghum(silage)"),
  N2O_long %>% filter(CropID == "Corn(grain)") %>% mutate(CropID = "Sorghum(grain)"),
  N2O_long %>% filter(CropID == "Canola") %>% mutate(CropID = "Camelina"),
  N2O_long %>% filter(CropID == "Wheat(spring)") %>% mutate(CropID = "Triticale")
) %>%
  arrange(ProvinceID)

#Without silage
write.csv(N2O_long, "../Output/Indirect N2O_FSN_Volatilization_Canada_GrainSplit.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean 0.0004901, median, 0.0003280

```

