---
title: "CA_N2O from CRN leaching"
author: "Dr. Chih-Yu Hung"
date: "2024-03-19"
output: html_document
---

## Prepare the data

For the leaching, we need the N from crop residue and P and PE. 

```{r Data, include=FALSE}
library(tidyverse)
SiteData<-read_rds("../Input/CA_Direct_Crop_2016-2022.rds") #SiteData
#SiteData<-read_rds("../Input/CA_Direct_Crop_2016-2022_WO_Silage.rds") # Without silage

#The last 5 years
SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5)) %>%
  select (c(RegionID:Fertilizer_Applied, P = Precipitation, PE = Evapo))


# For the Clean Fuel Regulation
# For the regulation and LCA model, AFOLU needs to provide a table of Leached N2O EF, which has EFs for provinceas (Atlantic) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Atlantic provinces. 
AT <- c("NL","NS","PE","NB")
SiteData <- SiteData %>%
  mutate (ProvinceID = case_when(         #combine the region
          Province_ID %in% AT ~ "AT",
          TRUE ~ Province_ID))
```

## Leaching factor  
The leaching factor is calcualted only based on the precipitation and potential evaporation. 
A linear regression was applied when the P/PE between 0.23-1


```{r Leaching factor}
source("Function for Indirect N2O EF_leaching.R") #The function 
CA_leaching <- N2OEF_leaching(SiteData)
N2O_long <- CA_leaching$Coef_Prov_Crop
```


## Add Sorghum, camelina, triticale data

Detail description can be found in the direction emission file or volatilization file


```{r adding data}
#Add a Crop column to Sorghum
Sorghum_rows <- N2O_long[N2O_long$CropID == "Corn", ]
Sorghum_rows$CropID <- "Sorghum"

# Add a Crop column to camelina
Camelina_rows <- N2O_long[N2O_long$CropID == "Canola", ]
Camelina_rows$CropID <- "Camelina"

# Add a Crop column to triticale
Triticale_rows <- N2O_long[N2O_long$CropID == "Wheat(spring)", ]
Triticale_rows$CropID <- "Triticale"


N2O_long <- rbind(N2O_long, Sorghum_rows, Camelina_rows, Triticale_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

#Save the output for CSED
#write.csv(N2O_long, "../Output/Indirect N2O_CRN_Leaching_Canada.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean 0.00151817, median, 0.0013491

#Without silage
#write.csv(N2O_long, "../Output/Indirect N2O_CRN_Leaching_Canada_WO_Silage.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean 0.0015764, median, 0.00135491


```
