---
title: "CA_N2O from FSN leaching"
author: "Dr. Chih-Yu Hung"
date: "2024-03-19"
output: html_document
---

## Prepare the data

For the leaching, we just need the `Fertilizer_Applied` (From Crop type and real Fertilizer_2016-2022.RData) and FracLeach and EF5 for each ecodistricts. The FracLeach is determined by precipitation and evapotranspiration (From INPUT-EF_Fixed_Para.csv), which is the 1981-2010 normals.  


## Add Sorghum, camelina, triticale data

Detail description can be found in the direction emission file or volatilization file

```{r Data, include=FALSE}
library(tidyverse)
source("Function for Indirect N2O EF_leaching.R") #The function 
SiteData_GrainAgg <- read_rds("../Input/CA_Direct_2016-2022_GrainAgg.rds") #SiteData
SiteData_GrainSplit <- read_rds("../Input/CA_Direct_2016-2022_GrainSplit.rds") #SiteData

# For the Clean Fuel Regulation
# For the regulation and LCA model, AFOLU needs to provide a table of Leached N2O EF, which has EFs for provinceas (Atlantic) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Atlantic provinces. 
AT <- c("NL","NS","PE","NB")

#The last 5 years
SiteData_GrainAgg <- SiteData_GrainAgg %>%
  filter (Year > (max(Year)-5)) %>%
  select (c(RegionID:Fertilizer_Applied, P = Precipitation, PE = Evapo)) %>%
  mutate (ProvinceID = case_when(         #combine the region
          Province_ID %in% AT ~ "AT",
          TRUE ~ Province_ID))


#The last 5 years
SiteData_GrainSplit <- SiteData_GrainSplit %>%
  filter (Year > (max(Year)-5)) %>%
  select (c(RegionID:Fertilizer_Applied, P = Precipitation, PE = Evapo)) %>%
  mutate (ProvinceID = case_when(         #combine the region
          Province_ID %in% AT ~ "AT",
          TRUE ~ Province_ID))
```

## Leaching factor  
The leaching factor is calcualted only based on the precipitation and potential evaporation. 
A linear regression was applied when the P/PE between 0.23-1


```{r Leaching factor}
#Grain Agg
CA_leaching <- N2OEF_leaching(SiteData_GrainAgg)

N2O_long <- CA_leaching$Coef_Prov_Crop

N2O_long <- bind_rows(
  N2O_long,
  N2O_long %>% filter(CropID == "Corn") %>% mutate(CropID = "Sorghum"),
  N2O_long %>% filter(CropID == "Canola") %>% mutate(CropID = "Camelina"),
  N2O_long %>% filter(CropID == "Wheat(spring)") %>% mutate(CropID = "Triticale")
) %>%
  arrange(ProvinceID)

#Save the output for CSED
#write.csv(N2O_long, "../Output/Indirect N2O_FSN_Leaching_Canada_GrainAgg.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean 0.001565, median, 0.001354

#Grain Split
CA_leaching <- N2OEF_leaching(SiteData_GrainSplit)

N2O_long <- CA_leaching$Coef_Prov_Crop

# Add surrogate crops
N2O_long <- bind_rows(
  N2O_long,
  N2O_long %>% filter(CropID == "Corn(silage)") %>% mutate(CropID = "Sorghum(silage)"),
  N2O_long %>% filter(CropID == "Corn(grain)") %>% mutate(CropID = "Sorghum(grain)"),
  N2O_long %>% filter(CropID == "Canola") %>% mutate(CropID = "Camelina"),
  N2O_long %>% filter(CropID == "Wheat(spring)") %>% mutate(CropID = "Triticale")
) %>%
  arrange(ProvinceID)

#Save the output for CSED
#write.csv(N2O_long, "../Output/Indirect N2O_FSN_Leaching_Canada_GrainSplit.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean 0.001556, median, 0.001342


```
