---
title: "CA_N2O from leaching"
author: "Dr. Chih-Yu Hung"
date: "2024-03-19"
output: html_document
---

## Prepare the data

For the leaching, we just need the `Fertilizer_Applied` (From Crop type and real Fertilizer_2016-2022.RData) and FracLeach and EF5 for each ecodistricts. The FracLeach is determined by precipitation and evapotranspiration (From INPUT-EF_Fixed_Para.csv), which is the 1981-2010 normals.  


```{r Data, include=FALSE}
library(tidyverse)
#Obtain fertilization amount data
load("../Input/Crop type and real Fertilizer_2016-2022.RData")#Fert_corp

#Merge grain and silage for corn and fall and spring for rye
Fert_crop <- Fert_crop %>%
  filter(CropID != "Sugarbeet") %>% #no longer need for this project
 # Harmonise the CropID
  mutate(
    CropID = str_replace_all(
      CropID,
      c(
        "Corn\\((?:silage|grain)\\)" = "Corn",
        "Rye\\((?:fall|spring)\\)"   = "Rye"
      )
    )
  ) %>% 
  group_by(Province_ID, Ecodistrict_ID, Inv_Year_ID, CropID) %>% 
  summarise(
    Crop_Real_N = sum(Crop_Real_N, na.rm = TRUE),
    across(-Crop_Real_N, first),
    .groups = "drop"
  )


#Covert the colnames
CA_leaching <- Fert_crop %>%
  select(RegionID,ProvinceID = Province_ID, Year = Inv_Year_ID, CropID, Ecodistrict_ID, Fertilizer_Applied = Crop_Real_N) %>%
  mutate(RegionID = paste(CropID, ProvinceID, Ecodistrict_ID, Year, sep = "_"))

#Save data
#save(CA_leaching, file = "../Input/CA_leaching.RData")

#Inner_join, assigning P and PE to Fertilizer_Applied
INPUT_Fixed <-read.csv("../Input/INPUT-EF_Fixed_Para.csv") #for Precipitation and potential evaporation data

SiteData <- CA_leaching %>%
  inner_join(INPUT_Fixed, by = c("ProvinceID" = "Province_ID", "Ecodistrict_ID" = "Ecodistrict_ID"))
  #There is no Ecodistrict 948 in INPUT_FIXED, 948 has very low fertilization though
  #cannot combine the following code as it different length casued error

SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))


# For the Clean Fuel Regulation
# For the regulation and LCA model, AFOLU needs to provide a table of Leached N2O EF, which has EFs for provinceas (Atlantic) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Atlantic provinces. 
AT <- c("NL","NS","PE","NB")
SiteData <- SiteData %>%
  mutate (ProvinceID = case_when(         #combine the region
          ProvinceID %in% AT ~ "AT",
          TRUE ~ ProvinceID))
```

## Leaching factor  
The leaching factor is calcualted only based on the precipitation and potential evaporation. 
A linear regression was applied when the P/PE between 0.23-1


```{r Leaching factor}
source("Function for Indirect N2O EF_leaching.R") #The function 
CA_leaching <- N2OEF_leaching(SiteData)
N2O_long <- CA_leaching$Coef_Prov_Crop
```


## Add Sorghum, camelina, triticale data

Detail description can be found in the direction emission file or volatilization file


```{r adding data}
#Add a Crop column to Sorghum
Sorghum_rows <- N2O_long[N2O_long$CropID == "Corn", ]
Sorghum_rows$CropID <- "Sorghum"

# Add a Crop column to camelina
Camelina_rows <- N2O_long[N2O_long$CropID == "Canola", ]
Camelina_rows$CropID <- "Camelina"

# Add a Crop column to triticale
Triticale_rows <- N2O_long[N2O_long$CropID == "Wheat(spring)", ]
Triticale_rows$CropID <- "Triticale"


N2O_long <- rbind(N2O_long, Sorghum_rows, Camelina_rows, Triticale_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

#Save the output for CSED
write.csv(N2O_long, "../Output/Indirect N2O_Leaching_Canada.csv", row.names = FALSE)
summary(N2O_long$N2O.IEF)
#mean 0.001565, median, 0.001354

```
