---
title: "N2O from leaching"
author: "Dr. Chih-Yu Hung"
date: "2024-03-19"
output: html_document
---

## Prepare the data

For the leaching, we just need the `Fertilizer_Applied` (From Crop type and real Fertilizer_2016-2022.RData) and FracLeach and EF5 for each ecodistricts. The FracLeach is determined by precipitation and evapotranspiration (From INPUT-EF_Fixed_Para.csv), which is the 1981-2010 normals.  


```{r Data, include=FALSE}
library(tidyverse)
#Obtain fertilization amount data
load("../Input/Crop type and real Fertilizer_2016-2022.RData")#Fert_corp

#Covert the colnames
CA_leaching <- Fert_crop %>%
  select(RegionID,ProvinceID = Province_ID, Year = Inv_Year_ID, CropID, Ecodistrict_ID, Fertilizer_Applied = Crop_Real_N)

#Save data
#save(CA_leaching, file = "Input/CA_leaching.RData")

#Inner_join, assigning P and PE to Fertilizer_Applied
INPUT_Fixed <-read.csv("../Input/INPUT-EF_Fixed_Para.csv") #for Precipitation and potential evaporation data

SiteData <- CA_leaching %>%
  inner_join(INPUT_Fixed, by = c("ProvinceID" = "Province_ID", "Ecodistrict_ID" = "Ecodistrict_ID"))
  #There is no Ecodistrict 948 in INPUT_FIXED, 948 has very low fertilization though
  #cannot combine the following code as it different length casued error

SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))

```

## Leaching factor  
The leaching factor is calcualted only based on the precipitation and potential evaporation. 
A linear regression was applied when the P/PE between 0.23-1


```{r Leaching factor}
source("Function for Indirect N2O EF_leaching.R") #The function 
CA_leaching <- N2OEF_leaching(SiteData)
```

## For the Clean Fuel Regulation

For the regulation and LCA model, AFOLU needs to provide a table of Leached N2O EF, which has EFs for provinceas (Maritime) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Maritime provinces. 

```{r CFR and LCA model}
#Atlantic provinces
AT <- c("NL","NS","PE","NB")

N2O_long<- CA_leaching[[4]] %>%
  mutate (Region = case_when(         #combine the region
          ProvinceID %in% AT ~ "AT",
          TRUE ~ProvinceID)) %>%
  group_by(Region,CropID) %>%
  summarise(N2O.IEF = ifelse(sum(Fertilizer_Applied) > 0, sum(N2O)/sum(Fertilizer_Applied), 0))
```

## Add Sorghum data
In Canada, we have very little amount of grain sorghum. Considering the close relationship between grain Sorghum and corn and their similar N requirement and fertilizer management plan. We are going to use N2O EF of corn for sorghum. 

Ref:
Jordan B., Fertilizing to help maximize grain sorghum yields. https://www.cropscience.bayer.us/articles/channel/fertilizing-to-help-maximize-grain-sorghum-yields. access 24 May, 2024.
Vanderlip, R., Roozeboom, K., Fjell, D., Shroyer, J., Kok, H., Regehr, D., Whitney, D., Rogers, D.H., Alam, M., Jardine, D. Brooks, H.L., Taylor, R.K., Harner III, J.P., and Langemeier, L.N. 1998. Grain sorghum production handbook. C-687 (Revised). Kansas State University. https://bookstore.ksre.ksu.edu/pubs/c687.pdf.
Haag, L. (In cooperation with: Diaz, D. R., Obour, A., and Schlegel, A.) 2018. Grain sorghum fertility management. Kansas State University. https://www.northwest.k-state.edu/agronomy/documents/presentation_pdfs/2018%20Grain%20Sorghum%20Fertility%20Management%20SorghumU%202spp.pdf.

```{r adding sorghum data}
corn_rows <- N2O_long[N2O_long$CropID == "Corn(grain)", ]
# Modify the Crop column to Sorghum
sorghum_rows <- corn_rows
sorghum_rows$CropID <- "Sorghum"
N2O_long <- rbind(N2O_long, sorghum_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

```

## Add Camelina data
In Canada, the optimal fertilization plan for camelina was not determined. Recommended application range is 60-100 kg N /ha. However, CFIA (2017) suggested using fertilization plan of canola if further information is not available. 

SK's fertilizer guide showed camelina is planted in brown soil and their default fertilizer rate is 75 kg N /ha in brown soils, which is lower than 85 kg N /ha for canola in brown soils in SK. Therefore we applied a correction factor 75/85 to the N2O EF in SK. (Agriculture Saskatchewan, 2024)


Other province didn't have fertilizer guide for cameline (e.g. Manitoba Agriculture), AFOLU therefore used EFs of canola.  


CFIA (2017) The Biology of Camelina sativa (L.) Crantz (Camelina). https://inspection.canada.ca/en/plant-varieties/plants-novel-traits/applicants/directive-94-08/biology-documents/camelina-sativa
Manitoba agriculture (2024) Soil Fertility Guide. https://www.gov.mb.ca/agriculture/crops/soil-fertility/soil-fertility-guide/general-fertilizer-recs-without-soil-test.html. Access 2024. 05. 15.
Agriculture Saskatchewan (2024) Crop planning guide online calculator 2024. 
https://publications.saskatchewan.ca/#/products/122662

```{r adding Camelina data}
canola_rows <- N2O_long[N2O_long$CropID == "Canola", ]
# Modify the Crop column to Sorghum
camelina_rows <- canola_rows
camelina_rows$CropID <- "Camelina"
#camelina_rows$N2O.IEF[camelina_rows$Region == "AB"] <- camelina_rows$N2O.IEF[camelina_rows$Region == "AB"] *0.9
camelina_rows$N2O.IEF[camelina_rows$Region == "SK"] <- camelina_rows$N2O.IEF[camelina_rows$Region == "SK"] * (75/85)

N2O_long <- rbind(N2O_long, camelina_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

```


## Add Triticale data
Only QC, ON, SK, MB, AB (QC, ON and MB has very small amount <10000 metric ton) have triticale production

Government of Alberta - Ministry of Agriculture and Irrigation (2004) recommends to use 35-90 (lb/acre) for triticale and 35-110 (lb/acre) for spring wheat on irrigated lands. AFOLU assumed avg. fertilization rate for triticale in AB is 62.5 lb/acre, which is 86% of fertilization rate of spring wheat. Since there is no documents to show the actual fertilizer usage on triticale production, AFOLU applied 86% as correction factor for triticale. 


Manitoba Agriculture (2024) recommends to use less fertilizer for triticale than wheat. The reduction rate from 0.67 to 0.73 depends on the previous crop, (e.g. fallow and stubble 0.67, and after grass 0.73). There is no information to know the previous crops for triticale field. A adjustment factor 0.7 is applied to triticale fertilization in Manitoba. 



Government of Alberta - Ministry of Agriculture and Irrigation (2004) Alberta Fertilizer Guide. https://open.alberta.ca/publications/2909624   
Agriculture Manitoba (2024) Triticale Production and Management
https://www.gov.mb.ca/agriculture/crops/crop-management/triticale.html access on June 10, 2024
Agriculture Manitoba (2024) Spring Wheat Production and Management. 
https://www.gov.mb.ca/agriculture/crops/crop-management/spring-wheat.html access on June 10, 2024

```{r adding Tritical data}
wheat_rows <- N2O_long[N2O_long$CropID == "Wheat(spring)", ]
# Modify the Crop column to Sorghum
triticale_rows <- wheat_rows
triticale_rows$CropID <- "Triticale"
triticale_rows$N2O.IEF[triticale_rows$Region == "AB"] <- triticale_rows$N2O.IEF[triticale_rows$Region == "AB"] *0.86
triticale_rows$N2O.IEF[triticale_rows$Region == "MB"] <- triticale_rows$N2O.IEF[triticale_rows$Region == "MB"] * 0.7

N2O_long <- rbind(N2O_long, triticale_rows)
N2O_long <- N2O_long[order(N2O_long$CropID),]

```




## Making table for report

``` {r pivot to wide and export table}

#Replace 0 to NA
N2O_long[N2O_long == 0] <- NA
                      
# Pivot N2O.IEF
N2O_wide <- N2O_long %>%
  pivot_wider(names_from =  Region, values_from = N2O.IEF) #kg N2O-N per kg N

#Save the output for CSED
write.csv(N2O_wide, "../Output/Indirect N2O_leaching_Canada.csv")


```
## Comparison 

1. Our CRF shows the average FRACLEACH is 0.177703366 
 (Table 3.D from CAN_2023_2021)
 
2. In Open LCA model, they have FracLEACH and EF5, the average FracLEACHis 0.2150 and EF5 is 0.0075 
  This means the average EF could be 0.0016127. (3. N2O Emission Factors)
  
3. In my calculation, the average IEF is 0.001421 (implied FRACLEACH = 0.1894667) This is lower than the two values above. (exclude sorghum)

```{r result}

tibble(N2O_wide, caption = "Average N2O IEF from leaching for crops in 2018-2022") 


#compare the original OpenLCA results
load("../Input/OpenLCA.RData")

#Inner join OpenLCA to N2O_long for comarios
Leaching_comp <- N2O_long %>%
  inner_join(OpenLCA, by = c ("Region" = "Region"))

Leaching_comp <- Leaching_comp %>%
  mutate(Diff = ifelse(N2O.IEF != 0, N2O.IEF- (FracLEACH*EF5), 0)) %>%
  mutate(Diff.p = round(Diff/(FracLEACH*EF5)*100,1))
#select data for pivot to wide, need to take p first, otherwise data was elminated from Leaching_comp
Leaching_comp.p <- Leaching_comp %>%
  select(Region, CropID, Diff.p)
Leaching_comp <- Leaching_comp %>%
  select(Region, CropID, Diff)


# Pivot Difference
Leaching_comp_wide <- Leaching_comp %>%
  pivot_wider(names_from =  Region, values_from = Diff) #kg N2O-N per kg N

Leaching_comp.p_wide <- Leaching_comp.p %>%
  pivot_wider(names_from =  Region, values_from = Diff.p) #kg N2O-N per kg N

```


## Calculate the N2O intensity for crop
Calculate the total N2O from leached N for every province and crops 
After that, divide the total N2O by crop production in the provinces
```{r N2O intensity for crop}
# load("../Input/Crop_production_prov.RData")
# N_leach_prov <- N_leach %>%
#   group_by(Province_ID,CropIII_ID) %>%
#   summarise(N_leach_prov = sum(N_leach))
# 
# N_intensity_leach_prov <-N_leach_prov %>%
#   inner_join(Production, by = c("Province_ID" = "Province",
#                                 "CropIII_ID" = "CropID")) %>%
#   mutate(N_intensity = ifelse(Production >0, N_leach_prov/(Production*10^3), 0))
```

## Calculate the MT

```{r Maritime province}
# MT <- c("NL","NS","PE","NB")
# N_intensity_leach_region <- N_intensity_leach_prov %>%
#   mutate (Region = case_when(
#           Province_ID %in% MT ~ "MT",
#           TRUE ~Province_ID)) %>%
#   group_by(Region,CropIII_ID,Crop) %>%
#   summarise(EF = round(mean(N_intensity),5))
# 
# 
# # Convert long table to wide table
# N_leach_result <- pivot_wider(N_intensity_leach_region,
#                               names_from = Region, values_from = EF)
# 
# write.csv(N_leach_result, "../Output/Indirect N2O_Leach.csv")
```