---
title: "N2O from leaching"
author: "Dr. Chih-Yu Hung"
date: "2024-03-19"
output: html_document
---

```{r Data, include=FALSE}
library(tidyverse)
INPUT_Fixed <-read.csv("../Input/INPUT-EF_Fixed_Para.csv") #for Precipitation and potential evaporation data
load("../Input/Crop type and real Fertilizer.RData")

```

## Leaching factor  
The leaching factor is calcualted only based on the precipitation and potential evaporation. 
A linear regression was applied when the P/PE between 0.23-1


```{r Leaching factor}
Frac_leach <- INPUT_Fixed %>%
  mutate(Frac_leach = ifelse(P >= PE,0.3,ifelse(P/PE <= 0.23,0.05,0.3247*(P/PE)-0.0247))) %>%
  mutate(Leach_factor = Frac_leach *0.0075) #EF5 0.0075 kg N2O-N kg-1 N
```

## Innerjoing leaching factor and real N fertilization for specific crop
We can calculate the real N leaching for specific crop at ecodistrict level

```{r Leaching factor}
N_leach <- Fert_crop %>%
  inner_join(Frac_leach, by = c("Province_ID" = "Province_ID",
                                "Ecodistrict_ID" = "Ecodistrict_ID")) %>%
  mutate(N_leach = Crop_Real_N * Leach_factor)
# N_leach = N2O from leached N (kg N2O-N)

```

## Calculate the N2O intensity for crop
Calculate the total N2O from leached N for every province and crops 
After that, divide the total N2O by crop production in the provinces
```{r N2O intensity for crop}
load("../Input/Crop_production_prov.RData")
N_leach_prov <- N_leach %>%
  group_by(Province_ID,CropIII_ID) %>%
  summarise(N_leach_prov = sum(N_leach))

N_intensity_leach_prov <-N_leach_prov %>%
  inner_join(Production, by = c("Province_ID" = "Province",
                                "CropIII_ID" = "CropID")) %>%
  mutate(N_intensity = ifelse(Production >0, N_leach_prov/(Production*10^3), 0))
```

## Calculate the MT

```{r Maritime province}
MT <- c("NL","NS","PE","NB")
N_intensity_leach_region <- N_intensity_leach_prov %>%
  mutate (Region = case_when(
          Province_ID %in% MT ~ "MT",
          TRUE ~Province_ID)) %>%
  group_by(Region,CropIII_ID,Crop) %>%
  summarise(EF = round(mean(N_intensity),5))


# Convert long table to wide table
N_leach_result <- pivot_wider(N_intensity_leach_region,
                              names_from = Region, values_from = EF)

write.csv(N_leach_result, "../Output/Indirect N2O_Leach.csv")

```