---
title: "N2O from leaching"
author: "Dr. Chih-Yu Hung"
date: "2024-03-19"
output: html_document
---

## Prepare the data

For the leaching, we just need the `Fertilizer_Applied` (From Crop type and real Fertilizer_2016-2022.RData) and FracLeach and EF5 for each ecodistricts. The FracLeach is determined by precipitation and evapotranspiration (From INPUT-EF_Fixed_Para.csv), which is the 1981-2010 normals.  

We need to inner_join them

```{r Data, include=FALSE}
library(tidyverse)
INPUT_Fixed <-read.csv("../Input/INPUT-EF_Fixed_Para.csv") #for Precipitation and potential evaporation data

#Obtain fertilization amount data
load("../Input/Crop type and real Fertilizer_2016-2022.RData")#Fert_corp
#Covert the colnames
CA_leaching <- Fert_crop %>%
  select(RegionID,ProvinceID = Province_ID, Year = Inv_Year_ID, CropID, Ecodistrict_ID, Fertilizer_Applied = Crop_Real_N)
#save(CA_leaching, file = "Input/CA_leaching.RData")


#Inner_join, assigning P and PE to Fertilizer_Applied

SiteData <- CA_leaching %>%
  inner_join(INPUT_Fixed, by = c("ProvinceID" = "Province_ID", "Ecodistrict_ID" = "Ecodistrict_ID"))
  #There is no Ecodistrict 948 in INPUT_FIXED, 948 has very low fertilization though
  #cannot combine the following code as it different length casued error

SiteData <- SiteData %>%
  filter (SiteData$Year > (max(SiteData$Year)-5))

```


## Prepare the data

For the leaching, we just need the `Fertilizer_Applied` (From Crop type and real Fertilizer_2016-2022.RData) and FracLeach and EF5 for each ecodistricts. The FracLeach is determined by precipitation and evapotranspiration (From INPUT-EF_Fixed_Para.csv), which is the 1981-2010 normals.  


## Leaching factor  
The leaching factor is calcualted only based on the precipitation and potential evaporation. 
A linear regression was applied when the P/PE between 0.23-1


```{r Leaching factor}
source("../Function for Indirect N2O EF_leaching.R") #The function 
CA_leaching <- N2OEF_leaching(SiteData)
```

## For the Clean Fuel Regulation

For the regulation and LCA model, AFOLU needs to provide a table of Leached N2O EF, which has EFs for provinceas (Maritime) and specific crops. AFOLU calculated weighted IEF based on fertilizer amount for Maritime provinces. 

```{r CFR and LCA model}
#Maritime provinces
MT <- c("NL","NS","PE","NB")

N2O_long<- CA_leaching[[4]] %>%
  mutate (Region = case_when(         #combine the region
          ProvinceID %in% MT ~ "MT",
          TRUE ~ProvinceID)) %>%
  group_by(Region,CropID) %>%
  summarise(N2O.IEF = ifelse(sum(Fertilizer_Applied) > 0, sum(N2O)/sum(Fertilizer_Applied), 0))

# Pivot N2O.IEF
N2O_wide <- N2O_long %>%
  pivot_wider(names_from =  Region, values_from = N2O.IEF) #kg N2O-N per kg N


#Save the output for CSED
write.csv(N2O_wide, "../Output/Indirect N2O_leaching_Canada.csv")

```


## Comparison 

1. Our CRF shows the average FRACLEACH is 0.177703366 
 (Table 3.D from CAN_2023_2021)
 
2. In Open LCA model, they have FracLEACH and EF5, the average FracLEACHis 0.2150 and EF5 is 0.0075 
  This means the average EF could be 0.0016127. (3. N2O Emission Factors)
  
3. In my calculation, the average IEF is 0.001421 (implied FRACLEACH = 0.1894667) This is lower than the two values above. 

```{r result}

tibble(N2O_wide, caption = "Average N2O IEF from leaching for crops in 2018-2022") 


#compare the original OpenLCA results
load("../Input/OpenLCA.RData")

#Inner join OpenLCA to N2O_long for comarios
Leaching_comp <- N2O_long %>%
  inner_join(OpenLCA, by = c ("Region" = "Region"))

Leaching_comp <- Leaching_comp %>%
  mutate(Diff = ifelse(N2O.IEF != 0, N2O.IEF- (FracLEACH*EF5), 0)) %>%
  mutate(Diff.p = round(Diff/(FracLEACH*EF5)*100,1))
#select data for pivot to wide, need to take p first, otherwise data was elminated from Leaching_comp
Leaching_comp.p <- Leaching_comp %>%
  select(Region, CropID, Diff.p)
Leaching_comp <- Leaching_comp %>%
  select(Region, CropID, Diff)


# Pivot Difference
Leaching_comp_wide <- Leaching_comp %>%
  pivot_wider(names_from =  Region, values_from = Diff) #kg N2O-N per kg N

Leaching_comp.p_wide <- Leaching_comp.p %>%
  pivot_wider(names_from =  Region, values_from = Diff.p) #kg N2O-N per kg N


```





## Calculate the N2O intensity for crop
Calculate the total N2O from leached N for every province and crops 
After that, divide the total N2O by crop production in the provinces
```{r N2O intensity for crop}
# load("../Input/Crop_production_prov.RData")
# N_leach_prov <- N_leach %>%
#   group_by(Province_ID,CropIII_ID) %>%
#   summarise(N_leach_prov = sum(N_leach))
# 
# N_intensity_leach_prov <-N_leach_prov %>%
#   inner_join(Production, by = c("Province_ID" = "Province",
#                                 "CropIII_ID" = "CropID")) %>%
#   mutate(N_intensity = ifelse(Production >0, N_leach_prov/(Production*10^3), 0))
```

## Calculate the MT

```{r Maritime province}
# MT <- c("NL","NS","PE","NB")
# N_intensity_leach_region <- N_intensity_leach_prov %>%
#   mutate (Region = case_when(
#           Province_ID %in% MT ~ "MT",
#           TRUE ~Province_ID)) %>%
#   group_by(Region,CropIII_ID,Crop) %>%
#   summarise(EF = round(mean(N_intensity),5))
# 
# 
# # Convert long table to wide table
# N_leach_result <- pivot_wider(N_intensity_leach_region,
#                               names_from = Region, values_from = EF)
# 
# write.csv(N_leach_result, "../Output/Indirect N2O_Leach.csv")

```