---
title: "US_N2O from leaching"
author: "Dr. Chih-Yu Hung"
date: "2025-02-24"
output: html_document
---

## Prepare the data



```{r Data, include=FALSE}
library(tidyverse)
#Obtain fertilization amount data
USData <- read_csv("C:/Users/HungC/OneDrive - EC-EC/2023/Fuel LCA/US data/Outputs/0.3_US_leaching.csv")

#CropID needs to combine CROP and CLASS 
# Apply the mapping to the dataset
source("../Crop mapping.R")
USData <- USData %>%
  mutate(CropID = crop_mapping(CROP, CLASS)) %>%
  filter(!is.na(CropID))  # Remove unwanted crops

#Prepare RegionID
USData <- USData %>%
  mutate(RegionID = str_c(CropID, STATE, COUNTY, YEAR, sep = "_"))


```

## Leaching factor  
The leaching factor is calculalted only based on the precipitation and potential evaporation. 
A linear regression was applied when the P/PE between 0.23-1


```{r Leaching factor}
source("Function for Indirect N2O EF_leaching.R") #The function 
US_leaching <- N2OEF_leaching(USData,
                              ProvinceCol = "STATE", FertCol = "Fert_LCA_kg" ,YearCol = "YEAR")
```

## For the Clean Fuel Regulation

For the regulation and LCA model, AFOLU needs to provide a table of Leached N2O EF, which has EFs for States  and specific crops. 

```{r CFR and LCA model}
# Clean and write the table for LCA model 
US_leaching_LCA <- US_leaching$Coef_Prov_Crop %>%
  select(ProvinceID,CropID,N2O.IEF) %>%
  rename(State = ProvinceID)
  
write.csv(US_leaching_LCA, "../Output/Indirect N2O_leaching_USA.csv", row.names = FALSE)  
  
```




## Making table for report

``` {r pivot to wide and export table}

# #Replace 0 to NA
# N2O_long[N2O_long == 0] <- NA
#                       
# # Pivot N2O.IEF
# N2O_wide <- N2O_long %>%
#   pivot_wider(names_from =  Region, values_from = N2O.IEF) #kg N2O-N per kg N
# 
# #Save the output for CSED
# write.csv(N2O_wide, "../Output/Indirect N2O_leaching_Canada.csv")
# 

```


## Comparison 

1. CA CRT shows the average FRACLEACH is 0.177703366 
 (Table 3.D from CAN_2023_2021)
 
2. In Open LCA model, they have FracLEACH and EF5, the average FracLEACHis 0.2150 and EF5 is 0.0075 
  This means the average EF could be 0.0016127. (3. N2O Emission Factors)
  
3. In my calculation, the average IEF is 0.001421 (implied FRACLEACH = 0.1894667) This is lower than the two values above. (exclude sorghum)

```{r result}

# tibble(N2O_wide, caption = "Average N2O IEF from leaching for crops in 2018-2022") 
# 
# 
# #compare the original OpenLCA results
# load("../Input/OpenLCA.RData")
# 
# #Inner join OpenLCA to N2O_long for comarios
# Leaching_comp <- N2O_long %>%
#   inner_join(OpenLCA, by = c ("Region" = "Region"))
# 
# Leaching_comp <- Leaching_comp %>%
#   mutate(Diff = ifelse(N2O.IEF != 0, N2O.IEF- (FracLEACH*EF5), 0)) %>%
#   mutate(Diff.p = round(Diff/(FracLEACH*EF5)*100,1))
# #select data for pivot to wide, need to take p first, otherwise data was elminated from Leaching_comp
# Leaching_comp.p <- Leaching_comp %>%
#   select(Region, CropID, Diff.p)
# Leaching_comp <- Leaching_comp %>%
#   select(Region, CropID, Diff)
# 
# 
# # Pivot Difference
# Leaching_comp_wide <- Leaching_comp %>%
#   pivot_wider(names_from =  Region, values_from = Diff) #kg N2O-N per kg N
# 
# Leaching_comp.p_wide <- Leaching_comp.p %>%
#   pivot_wider(names_from =  Region, values_from = Diff.p) #kg N2O-N per kg N

```

